---
title: Genetic differentiation on flowering time in Cerastium fontanum using a reciprocal transplant experiment
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 4
subtitle: Data preparation
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(tibble.width = Inf)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(knitr)
```

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
myPalette <- colorRampPalette(brewer.pal(11, "YlOrRd"))
```

# Data preparation

## Read data from Excel files

Data from mothers and fathers from the greenhouse data file

```{r}
data_mothers_fathers<-read_excel("C:/Users/avald/Dropbox/SU/Projects/cerastium_greenhouse/data/edited/Cerastium_greenhouse_spring_2022_editedAV.xlsx", 
                       sheet="extracted_data_mothers_fathers")
```

Phenology data from transplants

```{r}
data_phen_transplants<-read_excel("data/edited/Hengill_2022_Final.02_AV.xlsx", 
                       sheet="extracted_data_phen")
```

# IDs for matching with data from transplants

The numbers for mohters and fathers of the transplanted plants match mother and father in the greenhouse data when the plant came from a cross-pollination. When the plant came from a self-pollination, the number for the mother (=father) is the original seed mother ID (mother_originalID_iceland) without the plot number (numeric_code).

For example, for mother_originalID_iceland = H05 422:3/2, if this plant was selfed, the mother id would be 422.

There is also a "M3" mother - where does this come from? Is it M3 in Group 1 = H04 409:3/3? I suppose so by now.

Extract the numeric_code from mother_originalID_iceland

```{r}
data_mothers_fathers<-data_mothers_fathers%>%
  mutate(original_id_numeric=substr(sub(".* ", "",original_id_Iceland),1,3))
```

```{r}
unique(data_mothers_fathers$original_id_numeric)
```

Fix some that are not correct

```{r}
data_mothers_fathers<-data_mothers_fathers%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="HC2","871",original_id_numeric))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="24/","24",original_id_numeric_fixed))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="7:3","7",original_id_numeric_fixed))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="63:","63",original_id_numeric_fixed))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="48:","48",original_id_numeric_fixed))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="1:1","1",original_id_numeric_fixed))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="1:3","1",original_id_numeric_fixed))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="53:","53",original_id_numeric_fixed))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="99:","99",original_id_numeric_fixed))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="92:","92",original_id_numeric_fixed))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="110","1100",original_id_numeric_fixed))%>%
  mutate(original_id_numeric_fixed=ifelse(original_id_numeric=="114","1149",original_id_numeric_fixed))
```

Create new tibble with ids and temp

```{r}
data_temp_parents<-data_mothers_fathers%>%
  select(original_id_Iceland,new_id_malin,original_id_numeric_fixed,temp)
```

This tibble will be used to merge with transplant data, in order to get to know the temperature of the parents.
Merge based on new_id_malin for the crosspollinated plants, and on original_id_numeric_fixed for the selfpollinated plants.

# Phenology data prep

## Change mother noted as "M3"

There is one mother noted as M3. This should be M3 in group 1 = H04 409:3/3.
So I change this mother to 409.

```{r}
data_phen_transplants<-data_phen_transplants%>%
  mutate(mother=ifelse(mother=="M3",409,mother))
```

## Change column types

```{r}
data_phen_transplants<-data_phen_transplants%>%
  mutate_at(c("heat_zone","group","mother","father","id","repl","plot",
              "peak","fruit_coll","poll_type"),as.factor)%>%
  mutate_at(c("y","temp1","temp2","temp3"),as.numeric)%>%
  mutate_at(c("n_shoots"),as.integer)
```

## Calculate temperature

```{r}
data_phen_transplants<-data_phen_transplants%>%
  mutate(temp=ifelse(is.na(temp1)&is.na(temp2)&is.na(temp3),NA, 
                     # If all temps are NA --> temp = NA (4 cases)
                     ifelse(!is.na(temp1),temp1, 
                            # If temp1 is not NA --> temp = temp1
                            ifelse(is.na(temp1)&!is.na(temp2),temp2,
                                   # If temp1 is NA and temp 2 is not NA --> temp = temp2
                                   temp3)))) # Otherwise temp= temp3
```

## Create unique ids

```{r}
data_phen_transplants<-data_phen_transplants%>%
  mutate(unique_id=paste(heat_zone,id,repl,sep="_")) 
```

## Convert to long format and divide stage into real stage + grazing

"stage" column contains info on stage:

- V (Vegetative)	Only vegetative growth
- B1 (Bud stage 1)	Buds are just starting to form, very small
- B2 (Bud stage 2)	Buds are at medium size 
- B3 (Bud stage 3)	Buds are large but still completely closed
- B4 (Bud stage 4)	Buds are large and almost starting to flower
- F (Flowering)	At least one flower has opened
- F100 (100% flowering)	All flowers have opened, none are yet wilted
- W (Wilted)	At least one flower has wilted
- W100 (100% wilted)	All flowers are wilted

But also on grazing / lost and dead plants:

- X (Lost)	Neither plant nor nail was not found that day
- D (Dead)	Plant still in place but dead
- G (Grazed/gone)	Nail was found but no sign of plant
- SG (All stems grazed/gone)	All stems grazed/gone but with visible stubs
- /SG (Stem(s) grazed/gone from plant)	One or more stems have been grazed but there are still some left, phenology is measured for the one that is most mature
- A (Aborted)	All flowerbuds have stopped development, none have nor will open

Separate info from "stage" column into "stage" and "graz" columns. 

Remove plants that did not flower (where there is no record with stages F or F100 or W or W100). 

```{r}
data_phen_transplants_long<-full_join(
  data_phen_transplants%>%
    pivot_longer(cols=starts_with("week"),names_to="revision",
                 names_prefix="week",values_to="date")%>%
    select(unique_id,revision,date),
  data_phen_transplants%>%
    pivot_longer(cols=starts_with("stage"),names_to="revision",
                 names_prefix="stage",values_to="stage")%>%
    select(unique_id,plot,revision,stage))%>%
  separate(stage,into=c("stage","graz"),sep="/")

nrow(data_phen_transplants_long%>%distinct(unique_id)) # 858 ids


data_phen_transplants_long$graz <- with(data_phen_transplants_long,
                          ifelse(stage %in% c("X","G","D","SG","A"),
                            stage,graz)) # Info that should be on "graz" column
data_phen_transplants_long$stage  <- with(data_phen_transplants_long,
                          ifelse(stage %in% c("X","G","D","SG","A"),
                            NA,stage))

unique(data_phen_transplants_long$graz) # Correct
unique(data_phen_transplants_long$stage) # Correct

ids_fl <- filter(data_phen_transplants_long, stage %in% c("F","F100","W","W100"))%>%
  distinct(unique_id) 
nrow(ids_fl) # 603 ids that flowered
  
data_phen_transplants_long <- data_phen_transplants_long %>%
  filter(unique_id %in% as.vector(ids_fl$unique_id))

length(unique(data_phen_transplants_long$unique_id)) # Now including only the 603 ids that flowered
```

## Create column for vegetative / flowering / wilted

```{r}
data_phen_transplants_long <- data_phen_transplants_long %>%
  mutate(V_F=ifelse(stage %in% c("V","B1","B2","B3","B4"),"V",
                    ifelse(stage %in% c("F","F100","W","W100"),"F",NA)))%>%
  mutate(V_F_W=ifelse(stage %in% c("V","B1","B2","B3","B4"),"V",
                    ifelse(stage %in% c("F","F100"),"F",
                           ifelse(stage %in% c("W","W100"),"W",NA))))
```

## Calculate FFD (mean and corrected)

Get date for last recording with buds and first recording with flowers. Calculate FFD_meanDOY as the mean date between last recording with buds and first recording with flowers.

```{r message=FALSE, warning=FALSE}
# Calculate dates when the plant was first recorded as flowering
date_firstF  <- data_phen_transplants_long %>%
  group_by(unique_id) %>%
  filter(V_F=="F") %>% # Recorded as either flowering or wilted
  filter(date==min(date,na.rm=T)) %>%
  mutate(date_firstF=date) %>%
  dplyr::select(unique_id,date_firstF)

### CAREFUL FROM HERE ###

# Calculate median of the dates when the plant was flowering
date_medianF  <- data_phen_transplants_long %>%
  group_by(unique_id) %>%
  filter(V_F_W=="F") %>% # Recorded only as flowering?
  # If it was wilted on the first recording where it was not vegetative, 
  # we cannot calculate MFD
  filter(date==median(date,na.rm=T)) %>%
  mutate(date_medianF=date) %>%
  dplyr::select(unique_id,date_medianF)

# Calculate dates when the plant was last recorded as flowering
date_lastF  <- data_phen_transplants_long %>%
  group_by(unique_id) %>%
  filter(V_F_W=="F") %>% # Recorded only as flowering?
  # If it was wilted on the first recording where it was not vegetative, 
  # we cannot calculate LFD
  filter(date==max(date,na.rm=T)) %>%
  mutate(date_lastF=date) %>%
  dplyr::select(unique_id,date_lastF)

### TO HERE ###

# Calculate date when the plant was last recorded as vegetative
# (before being recorded as flowering)
date_lastV  <- data_phen_transplants_long %>%
  group_by(unique_id) %>%
  filter(V_F=="V") %>%
  full_join(date_firstF) %>%
  filter(date<=date_firstF) %>% 
  # Remove records after first record with flowers
  filter(date==max(date)) %>%
  mutate(date_lastV=date) %>%
  dplyr::select(unique_id,date_lastV)

data_phen_transplants_FFD <- data_phen_transplants_long %>%
  dplyr::select(unique_id,date,stage,V_F) %>%
  filter(!is.na(V_F)) %>%
  left_join(date_firstF) %>%
  filter(date<=date_firstF) %>% 
  # Remove records after first record with flowers
  left_join(date_lastV) %>%
  filter(date>=date_lastV|is.na(date_lastV)) %>%
  # Remove records before last record without flowers
  pivot_wider(id_cols=unique_id,names_from=V_F,
              values_from=c(stage,date_firstF,date_lastV)) %>%
  dplyr::select(unique_id,stage_V,stage_F,date_firstF_F,date_lastV_V) %>%
  rename(date_firstF=date_firstF_F,date_lastV=date_lastV_V) %>%
  ungroup() %>%
  arrange(unique_id) %>%
  mutate(date_firstF=yday(date_firstF),date_lastV=yday(date_lastV)) %>%
  # Change to julian dates
  mutate(FFD_meanDOY=ifelse(is.na(date_lastV),date_firstF-3.61, 
                            # Calculate first estimate of FFD:
                            # For those flowering in first revision
                            # assigned FFD as half a mean interval
                            # date_firstF-(7.22/2)=date_firstF-3.61
                            (date_lastV+date_firstF)/2)) %>%
  # For others, assigned FFD_meanDOY as mean of date_lastV and date_firstF
  mutate(interval=date_firstF-date_lastV) %>%
  # Interval between 5 and 11 days (mean=7.22 days)
  mutate(interval=ifelse(is.na(date_lastV),7.22,interval))%>%
  # Assign mean interval to those flowering in first revision (i.e. without date_lastV)
  mutate(term_lastV=ifelse(stage_V=="V"|stage_V=="B1",0.2*interval, 
                    ifelse(stage_V=="B2",0*interval,
                    ifelse(stage_V=="B3"|stage_V=="B4",-0.2*interval,NA))),
         term_firstF=ifelse(stage_F=="F"|stage_F=="F100",0.2*interval,
                     ifelse(stage_F=="W"|stage_F=="W100",-0.2*interval,NA))) %>% 
  # Calculate terms to add for correcting FFD
  rowwise() %>% 
  mutate(FFD_corr=sum(FFD_meanDOY,term_lastV,term_firstF,na.rm=T))
  # Calculate corrected FFD: FFD_meanDOY + both terms
```

## TO DO: Calculate MFD and LFD (mean and corrected)

# Create datafile

```{r}
data_transplants<-full_join(
  data_phen_transplants%>%select(unique_id,heat_zone,group,mother,father,id,
                                 repl,plot,x,y,peak,fruit_coll,comments1,comments2,
                                 poll_type,date_planted,temp),
  data_phen_transplants_FFD%>%select(unique_id,FFD_corr)
  ) # Full join, will assign NA to plants that did not flower (i.e. do not have a FFD_corr
```

# Merge with temp parents

This tibble will be used to merge with transplant data, in order to get to know the temperature of the parents.
Merge based on new_id_malin for the crosspollinated plants, and on original_id_numeric_fixed for the selfpollinated plants.

```{r}
data_temp_parents<-data_temp_parents%>%
  mutate(new_id_malin=as.factor(new_id_malin),
         original_id_numeric_fixed=as.factor(original_id_numeric_fixed))
```

```{r}
data_temp_mothers<-data_temp_parents%>%
  rename(temp_mother=temp)
data_temp_fathers<-data_temp_parents%>%
  rename(temp_father=temp)
```


```{r}
data_transplants<-rbind(
  # Crosspollinated
  data_transplants%>%filter(poll_type=="Crosspollinated")%>%
  left_join(data_temp_mothers%>%
              select(new_id_malin,temp_mother),
            join_by(mother == new_id_malin),
            multiple="first")%>% 
  # Because sometimes there are two records of the same mother/father,
  # but their temperature is the same
  left_join(data_temp_fathers%>%
              select(new_id_malin,temp_father),
            join_by(father == new_id_malin),
            multiple="first"),
  # Selfpollinated
  data_transplants%>%filter(poll_type=="Selfpollinated")%>%
  left_join(data_temp_mothers%>%
              select(original_id_numeric_fixed,temp_mother),
            join_by(mother == original_id_numeric_fixed),
            multiple="first")%>%
  left_join(data_temp_fathers%>%
              select(original_id_numeric_fixed,temp_father),
            join_by(father == original_id_numeric_fixed),
            multiple="first") 
)

data_transplants<-data_transplants%>%
  mutate(mother=as.character(mother),father=as.character(father))%>%
  mutate(father=ifelse(poll_type=="Selfpollinated",mother,father))%>%
  mutate(mother=as.factor(mother),father=as.factor(father))%>%
  mutate(temp_father=ifelse(poll_type=="Selfpollinated",temp_mother,temp_father))%>%
  # So far, selfed progeny have father=mother, and temp_father=temp_mother
  mutate(crossing=paste(mother,father,sep="_"))
```

# TO DO: Add info on peak

# TO DO: add info on shoot growth

# Save clean data as .csv

```{r}
write_csv(data_transplants,"data/clean/data_transplants.csv")
```

# Session info

```{r}
sessionInfo()
```









