---
title: Local adaptation in Cerastium fontanum using a reciprocal transplant experiment
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: 4
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
subtitle: Analyses with FFD
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(tibble.width = Inf)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(knitr)
library(ggthemes)
library(glmmTMB)
library(lme4)
library(lmerTest)
library(car)
library(sjPlot)
library(ggeffects)
library(performance)
library(DHARMa)
library(MASS)
```

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
myPalette <- colorRampPalette(brewer.pal(11, "YlOrRd"))
```

# Read clean data from .csv files

```{r}
data_transplants <- read_csv("data/clean/data_transplants.csv")
```

# Calculate temperature difference

Calculate absolute temperature difference between experienced temperature (temp) and mid-parental temperature.

```{r}
data_transplants<-data_transplants%>%
  mutate(temp_diff=abs(temp-((temp_mother+temp_father)/2)))
hist(data_transplants$temp_diff)
```


# Survival

Need to get data from file "Cerastium survival" --> for 2023

```{r}
data_transplants<-data_transplants%>%
  mutate(surv_22=ifelse(is.na(F_NF_A),0,1))
table(data_transplants$surv_22)
hist(data_transplants$surv_22)
```

Of 858 plants, 808 survived in 2022.

Model including plot as random, poll_type was not included as random because there are only 2 levels (and binomial glm showed no effect of poll_type on surv_22).

```{r}
mod_surv_22<-glmmTMB(surv_22~temp+temp_diff+(1|plot),data_transplants,
                     family="binomial")
# Using glmer, temp_diff significant, but model convergence warning.
# Adding glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
# also gives temp_diff NS, as glmmTMB.
summary(mod_surv_22)
```

## Predictions

```{r}
plot(ggpredict(mod_surv_22,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_surv_22,terms=c("temp_diff[all]")),add.data=T) # NS
```

## Diagnostics

```{r}
plot(simulateResiduals(mod_surv_22))
```

# Median height of flowering stem

```{r}
hist(data_transplants$median_h)
hist(log(data_transplants$median_h))
hist(sqrt(data_transplants$median_h))
```

```{r}
data_transplants<-data_transplants%>%
  mutate(median_h_log=log(median_h),median_h_sqrt=sqrt(median_h))
```

Model including plot as random, poll_type was not included as random because there are only 2 levels (and lm showed no effect of poll_type on median_h).

```{r}
mod_median_h_22<-glmmTMB(median_h~temp+temp_diff+(1|plot),data_transplants)
mod_median_h_log_22<-glmmTMB(median_h_log~temp+temp_diff+(1|plot),
                             data_transplants) # temp* (P= 0.0483)
mod_median_h_sqrt_22<-glmmTMB(median_h_sqrt~temp+temp_diff+(1|plot),
                              data_transplants)
summary(mod_median_h_22)
summary(mod_median_h_log_22)
summary(mod_median_h_sqrt_22)
```

## Predictions

```{r}
plot(ggpredict(mod_median_h_log_22,terms=c("temp[all]")),add.data=T)
```

## Diagnostics

```{r}
plot(simulateResiduals(mod_median_h_22))
plot(simulateResiduals(mod_median_h_log_22))
plot(simulateResiduals(mod_median_h_sqrt_22))
```

```{r}
AIC(mod_median_h_22,mod_median_h_log_22,mod_median_h_sqrt_22)
```

Keep model with log?

# Probability of flowering

```{r}
data_transplants<-data_transplants%>%
  mutate(flowering_22=ifelse(F_NF_A=="F",1,ifelse(F_NF_A=="NF",0,NA)))
table(data_transplants$flowering_22)
hist(data_transplants$flowering_22)
```
 
Of 808 plants that survived in 2022, 603 flowered and 204 did not flower (and 1 was aborted, not used here).
4 were grazed before they peaked, they are included as flowering here.

Does pollination type have an effect on probability of flowering?

```{r}
summary(glm(flowering_22~poll_type,data_transplants,family="binomial"))
plot(ggpredict(glm(flowering_22~poll_type,data_transplants,family="binomial")))
```

Yes! Cross-pollinated have a higher probability of flowering than self-pollinated.

Include in the model as random (but only 2 levels!) or fixed:

Model including plot as random, and poll_type as random:

```{r}
mod_flowering_22a<-glmmTMB(flowering_22~temp+temp_diff+(1|plot)+(1|poll_type),
                          data_transplants,family="binomial")
summary(mod_flowering_22a)
```
 
Model including plot as random, and poll_type as fixed:

```{r}
mod_flowering_22b<-glmmTMB(flowering_22~temp+temp_diff+poll_type+(1|plot),
                          data_transplants,family="binomial")
summary(mod_flowering_22b)
```

## Predictions

```{r}
plot(ggpredict(mod_flowering_22a,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_flowering_22b,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_flowering_22b,terms=c("poll_type")))
```

## Diagnostics

```{r}
plot(simulateResiduals(mod_flowering_22a))
plot(simulateResiduals(mod_flowering_22b))
```

 
# Number of flowers

## tot_fl_bodies

### Without median_h

Including only plants that flowered. 4 that were grazed before they peaked are not included

```{r}
hist((data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0))$tot_fl_bodies)
```

Does pollination type have an effect on number of flowers?

```{r}
summary(glm.nb(tot_fl_bodies~poll_type,
            data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0)))
plot(ggpredict(glm.nb(tot_fl_bodies~poll_type,
            data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0))))
```

Yes! Cross-pollinated have a higher number of flowers than self-pollinated.

Include in the model as random (but only 2 levels!) or fixed:

Model including plot as random, and poll_type as random:

```{r}
mod_nfl_22a<-glmmTMB(tot_fl_bodies~temp+temp_diff+(1|plot)+(1|poll_type),
                     data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_nfl_22a)
``` 

Model including plot as random, and poll_type as fixed:

```{r}
mod_nfl_22b<-glmmTMB(tot_fl_bodies~temp+temp_diff+poll_type+(1|plot),
                     data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_nfl_22b)
``` 

#### Predictions

```{r}
plot(ggpredict(mod_nfl_22a,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nfl_22b,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nfl_22b,terms=c("poll_type")))
```

#### Diagnostics

```{r}
plot(simulateResiduals(mod_nfl_22a))
plot(simulateResiduals(mod_nfl_22b))
testDispersion(simulateResiduals(mod_nfl_22a),alternative="greater")
testDispersion(simulateResiduals(mod_nfl_22b),alternative="greater")
```

### With median_h

Model including plot as random, and poll_type as random:

```{r}
mod_nfl_mh_22a<-glmmTMB(tot_fl_bodies~temp+temp_diff+median_h+
                          (1|plot)+(1|poll_type),
                        data_transplants%>%
                          filter(flowering_22==1&tot_fl_bodies>0),
                        family="nbinom2") # Poisson was overdispersed
summary(mod_nfl_mh_22a)
``` 

Model including plot as random, and poll_type as fixed:

```{r}
mod_nfl_mh_22b<-glmmTMB(tot_fl_bodies~temp+temp_diff+median_h+poll_type+
                          (1|plot),
                     data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_nfl_mh_22b)
``` 

#### Predictions

```{r}
plot(ggpredict(mod_nfl_mh_22a,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nfl_mh_22a,terms=c("median_h[all]")),add.data=T)
plot(ggpredict(mod_nfl_mh_22b,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nfl_mh_22b,terms=c("median_h[all]")),add.data=T)
plot(ggpredict(mod_nfl_mh_22b,terms=c("poll_type")))
```

#### Diagnostics

```{r}
plot(simulateResiduals(mod_nfl_mh_22a))
plot(simulateResiduals(mod_nfl_mh_22b))
testDispersion(simulateResiduals(mod_nfl_mh_22a),alternative="greater")
# Slightly overdispersed? p=0.048
testDispersion(simulateResiduals(mod_nfl_mh_22b),alternative="greater")
# Slightly overdispersed? p=0.048
```

## tot_fl_bodies_corr

Correct tot_fl_bodies to tot_fl_bodies_corr, to be at least as high as n_tot_fr.

```{r}
data_transplants<-data_transplants%>%
  mutate(tot_fl_bodies_corr=ifelse(n_tot_fr>tot_fl_bodies,n_tot_fr,
                                   tot_fl_bodies),
         diff_fr_fl=n_tot_fr-tot_fl_bodies)
```

Histogram of differences>

```{r}
hist(data_transplants$diff_fr_fl)
```

Negative differences: more flowers than fruits - OK
Positive differences: more fruits than flowers - not OK --> corrected

### Without median_h

Model including plot as random, and poll_type as random:

```{r}
mod_nflcorr_22a<-glmmTMB(tot_fl_bodies_corr~temp+temp_diff+(1|plot)+
                           (1|poll_type),
                     data_transplants%>%
                       filter(flowering_22==1&tot_fl_bodies_corr>0),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_nflcorr_22a)
``` 

Model including plot as random, and poll_type as fixed:

```{r}
mod_nflcorr_22b<-glmmTMB(tot_fl_bodies_corr~temp+temp_diff+poll_type+(1|plot),
                     data_transplants%>%
                       filter(flowering_22==1&tot_fl_bodies_corr>0),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_nflcorr_22b)
``` 

#### Predictions

```{r}
plot(ggpredict(mod_nflcorr_22a,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nflcorr_22b,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nflcorr_22b,terms=c("poll_type")))
```

#### Diagnostics

```{r}
plot(simulateResiduals(mod_nflcorr_22a))
plot(simulateResiduals(mod_nflcorr_22b))
testDispersion(simulateResiduals(mod_nflcorr_22a),alternative="greater")
testDispersion(simulateResiduals(mod_nflcorr_22b),alternative="greater")
```

### With median_h

Model including plot as random, and poll_type as random:

```{r}
mod_nflcorr_mh_22a<-glmmTMB(tot_fl_bodies_corr~temp+temp_diff+median_h+
                          (1|plot)+(1|poll_type),
                        data_transplants%>%
                          filter(flowering_22==1&tot_fl_bodies_corr>0),
                        family="nbinom2") # Poisson was overdispersed
summary(mod_nflcorr_mh_22a)
``` 

Model including plot as random, and poll_type as fixed:

```{r}
mod_nflcorr_mh_22b<-glmmTMB(tot_fl_bodies_corr~temp+temp_diff+median_h+
                              poll_type+(1|plot),
                     data_transplants%>%
                       filter(flowering_22==1&tot_fl_bodies_corr>0),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_nflcorr_mh_22b)
``` 

#### Predictions

```{r}
plot(ggpredict(mod_nflcorr_mh_22a,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nflcorr_mh_22a,terms=c("median_h[all]")),add.data=T)
plot(ggpredict(mod_nflcorr_mh_22b,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nflcorr_mh_22b,terms=c("median_h[all]")),add.data=T)
plot(ggpredict(mod_nflcorr_mh_22b,terms=c("poll_type")))
```

#### Diagnostics

```{r}
plot(simulateResiduals(mod_nflcorr_mh_22a))
plot(simulateResiduals(mod_nflcorr_mh_22b))
testDispersion(simulateResiduals(mod_nflcorr_mh_22a),alternative="greater")
# Slightly overdispersed? p=0.032
testDispersion(simulateResiduals(mod_nflcorr_mh_22b),alternative="greater")
# Slightly overdispersed? p=0.048
```

# Probability of fruiting

```{r}
data_transplants<-data_transplants%>%
  mutate(fruiting_22=ifelse(is.na(F_NF_A),NA, 
                            # NA for plants that did not survive
                            ifelse(F_NF_A=="A"|F_NF_A=="NF",NA, 
                                   # NA for plants that aborted 
                                   # or did not flower
                                   ifelse(n_count_fr>0,1,0)))) 
# 1/0 for plants that produced / did not produce fruits
table(data_transplants$fruiting_22)
hist(data_transplants$fruiting_22)
```

Of 603 plants that flowered in 2022, 582 produced fruits and 21 did not produce fruits.

Model including plot as random, poll_type was not included as random because there are only 2 levels (and binomial glm showed no effect of poll_type on fruiting_22).

```{r}
mod_fruiting_22<-glmmTMB(fruiting_22~temp+temp_diff+(1|plot),data_transplants,
                     family="binomial")
# Using glmer, temp_diff significant, but model convergence warning.
# Adding glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
# also gives temp_diff NS, as glmmTMB.
summary(mod_fruiting_22)
```

## Diagnostics

```{r}
plot(simulateResiduals(mod_fruiting_22))
```


# Number of fruits

Including only plants that flowered. 4 that were grazed before they peaked are not included

```{r}
hist((data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0))$n_tot_fr)
```

Does pollination type have an effect on number of fruits?

```{r}
summary(glm.nb(n_tot_fr~poll_type,
            data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0)))
plot(ggpredict(glm.nb(n_tot_fr~poll_type,
            data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0))))
```

Yes! Cross-pollinated have a higher number of fruits than self-pollinated.

Include in the model as random (but only 2 levels!) or fixed:

Model including plot as random, and poll_type as random:

```{r}
mod_nfr_22a<-glmmTMB(n_tot_fr~temp+temp_diff+(1|plot)+(1|poll_type),
                     data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_nfr_22a)
``` 

 Model including plot as random, and poll_type as fixed:

```{r}
mod_nfr_22b<-glmmTMB(n_tot_fr~temp+temp_diff+poll_type+(1|plot),
                     data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_nfr_22b)
``` 

### Predictions

```{r}
plot(ggpredict(mod_nfr_22a,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nfr_22b,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nfr_22b,terms=c("poll_type")))
```

### Diagnostics

```{r}
plot(simulateResiduals(mod_nfr_22a))
plot(simulateResiduals(mod_nfr_22b))
testDispersion(simulateResiduals(mod_nfr_22a),alternative="greater")
# Slightly overdispersed? p=0.044
testDispersion(simulateResiduals(mod_nfr_22b),alternative="greater")
# Slightly overdispersed? p=0.04
```

# Fruit set

```{r}
data_transplants<-data_transplants%>%
  mutate(fr_set_22=ifelse(fruiting_22==0,NA,n_tot_fr/tot_fl_bodies_corr))
hist(data_transplants$fr_set_22)
```

Does pollination type have an effect on fruit set?

```{r}
summary(glm(fr_set_22~poll_type,
            data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0),
            weights=tot_fl_bodies_corr,
            family="binomial"))
plot(ggpredict(glm(fr_set_22~poll_type,
            data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0),
            weights=tot_fl_bodies_corr,
            family="binomial")))
```

Yes! Self-pollinated have a higher fruit set than cross-pollinated (WHY??).

Include in the model as random (but only 2 levels!) or fixed:

Model including plot as random, and poll_type as random:

```{r}
mod_frset_22a<-glmmTMB(fr_set_22~temp+temp_diff+(1|plot)+(1|poll_type),
                       weights=tot_fl_bodies_corr,
                       data_transplants%>%
                         filter(flowering_22==1&tot_fl_bodies>0),
                       family="binomial")
summary(mod_frset_22a)
``` 

Model including plot as random, and poll_type as fixed:

```{r}
mod_frset_22b<-glmmTMB(fr_set_22~temp+temp_diff+poll_type+(1|plot),
                       weights=tot_fl_bodies_corr,
                       data_transplants%>%
                         filter(flowering_22==1&tot_fl_bodies>0),
                       family="binomial")
summary(mod_frset_22b)
``` 

#### Predictions

```{r}
plot(ggpredict(mod_frset_22a,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_frset_22a,terms=c("temp_diff[all]")),add.data=T)
plot(ggpredict(mod_frset_22b,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_frset_22b,terms=c("temp_diff[all]")),add.data=T)
plot(ggpredict(mod_frset_22b,terms=c("poll_type")))
```

#### Diagnostics

```{r}
plot(simulateResiduals(mod_frset_22a))
plot(simulateResiduals(mod_frset_22b))
```

# Number of seeds

Including only plants that produced fruits.

```{r}
hist((data_transplants%>%filter(fruiting_22==1))$n_tot_seed)
```

Does pollination type have an effect on number of flowers?

```{r}
summary(glm.nb(n_tot_seed~poll_type,
            data_transplants%>%filter(fruiting_22==1)))
plot(ggpredict(glm.nb(n_tot_seed~poll_type,
            data_transplants%>%filter(fruiting_22==1))))
```

Yes! Self-pollinated have a higher number of seeds than cross-pollinated (WHY??).

Include in the model as random (but only 2 levels!) or fixed:

Model including plot as random, and poll_type as random:

```{r}
mod_nseed_22a<-glmmTMB(n_tot_seed~temp+temp_diff+(1|plot)+(1|poll_type),
                     data_transplants%>%filter(fruiting_22==1),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_nseed_22a)
``` 

 Model including plot as random, and poll_type as fixed:

```{r}
mod_nseed_22b<-glmmTMB(n_tot_seed~temp+temp_diff+poll_type+(1|plot),
                     data_transplants%>%filter(fruiting_22==1),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_nseed_22b)
``` 

### Predictions

```{r}
plot(ggpredict(mod_nseed_22a,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nseed_22b,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_nseed_22b,terms=c("poll_type")))
```

### Diagnostics

```{r}
plot(simulateResiduals(mod_nseed_22a))
plot(simulateResiduals(mod_nseed_22b))
testDispersion(simulateResiduals(mod_nseed_22a),alternative="greater")
testDispersion(simulateResiduals(mod_nseed_22b),alternative="greater")
```

# Seeds per flower

```{r}
data_transplants<-data_transplants%>%
  mutate(n_seeds_per_fl_22=ifelse(fruiting_22==0,NA,
                                  n_tot_seed/tot_fl_bodies_corr))
hist(data_transplants$n_seeds_per_fl_22)
```

Does pollination type have an effect on number of seeds per flower?

```{r}
summary(glm.nb(n_seeds_per_fl_22~poll_type,
            data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0)))
plot(ggpredict(glm.nb(n_seeds_per_fl_22~poll_type,
            data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0))))
```

Yes! Self-pollinated have a higher number of seeds per flower than cross-pollinated (WHY??).

Include in the model as random (but only 2 levels!) or fixed:

Model including plot as random, and poll_type as random:

```{r}
mod_seedsperfl_22a<-glmmTMB(n_seeds_per_fl_22~temp+temp_diff+(1|plot)+
                              (1|poll_type),
                     data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_seedsperfl_22a)
``` 

 Model including plot as random, and poll_type as fixed:

```{r}
mod_seedsperfl_22b<-glmmTMB(n_seeds_per_fl_22~temp+temp_diff+poll_type+(1|plot),
                     data_transplants%>%filter(flowering_22==1&tot_fl_bodies>0),
                     family="nbinom2") # Poisson was overdispersed
summary(mod_seedsperfl_22b)
``` 

### Predictions

```{r}
plot(ggpredict(mod_seedsperfl_22a,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_seedsperfl_22b,terms=c("temp[all]")),add.data=T)
plot(ggpredict(mod_seedsperfl_22b,terms=c("poll_type")))
```

### Diagnostics

```{r}
plot(simulateResiduals(mod_seedsperfl_22a))
plot(simulateResiduals(mod_seedsperfl_22b))
testDispersion(simulateResiduals(mod_seedsperfl_22a),alternative="greater")
testDispersion(simulateResiduals(mod_seedsperfl_22b),alternative="greater")
```

# Session info

```{r}
sessionInfo()
```
