---
title: 'Analyses Cerastium transplants: Local adaptation'
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    toc: true
    toc_depth: '4'
  html_document:
    toc: true
    toc_depth: '4'
    df_print: paged
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 4
subtitle: Analyses with FFD
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(tibble.width = Inf)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Load libraries

```{r load packages, include=FALSE}
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(knitr)
library(ggthemes)
library(glmmTMB)
library(lme4)
library(lmerTest)
library(car)
library(sjPlot)
library(ggeffects)
library(performance)
library(DHARMa)
library(MASS)
library(grid)
library(gridExtra)
library(viridis)
library(aster)
library(ggpubr)
library(piecewiseSEM)
```

# Define ggplot themes

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
myPalette <- colorRampPalette(brewer.pal(11, "YlOrRd"))
```

# Read clean data from .csv files

```{r}
data_transplants <- read_csv("data/clean/data_transplants.csv")%>%
  mutate_at(c("unique_id","heat_zone","group","mother","father","id","repl",
              "plot","peak","fruit_coll","poll_type","crossing","F_NF_A",
              "year"),as.factor)%>%
  mutate_at(c("n_fl_stems","n_opened_fl","n_closed_buds","tot_fl_bodies",
              "median_h","per_invert_herb","n_stems_grazed","n_closed_coll_fr",
              "n_open_coll_fr","n_count_fr","n_tot_fr","n_seed_fr1",
              "n_seed_fr2","n_seed_fr3","n_seed_fr4","rest_combined",
              "n_tot_seed","tot_fl_bodies_corr"),as.integer)
```

# Check that temperatures for sires and dams coincide with temperatures in list of sires and dams:

```{r}
temps_dams_sires<-data_transplants%>%
  dplyr::select(mother,father,temp_mother,temp_father)%>%
  distinct()%>% # Removes duplicate rows
  pivot_longer(cols=starts_with("temp"),names_to="type",values_to="temp")%>%
  mutate(id = case_when(type=="temp_mother"~paste0("D",mother),
                        type=="temp_father"~paste0("S",father)))%>%
  dplyr::select(id,temp)
# Read csv file with temperatures from sire and dam list
temps_dams_sires_list<-read_csv("C:/Users/alici/Dropbox/SU/Projects/cerastium_greenhouse/data/clean/temps_dams_sires_list.csv")
# Merge the data frames on the "id" column
merged_data<-temps_dams_sires%>%
  inner_join(temps_dams_sires_list,by="id",suffix=c("_data","_list"))
# Check for discrepancies
discrepancies<-merged_data%>%filter(temp_data!=temp_list)
# View the discrepancies
print(discrepancies)  # Two discrepancies
```

Correct discrepancies:

Temperature for mother 24 should be 14.9, not 20.
Temperature for father 320 should be 9.0, not 13.6.

```{r}
data_transplants <- data_transplants %>%
  mutate(temp_mother = if_else(mother == 24, 14.9, temp_mother),
         temp_father = if_else(father == 320, 9.0, temp_father))
```

Check again for discrepancies:

```{r}
temps_dams_sires<-data_transplants%>%
  dplyr::select(mother,father,temp_mother,temp_father)%>%
  distinct()%>% # Removes duplicate rows
  pivot_longer(cols=starts_with("temp"),names_to="type",values_to="temp")%>%
  mutate(id = case_when(type=="temp_mother"~paste0("D",mother),
                        type=="temp_father"~paste0("S",father)))%>%
  dplyr::select(id,temp)
# Read csv file with temperatures from sire and dam list
merged_data<-temps_dams_sires%>%
  inner_join(temps_dams_sires_list,by="id",suffix=c("_data","_list"))
# Check for discrepancies
discrepancies<-merged_data%>%filter(temp_data!=temp_list)
# View the discrepancies
print(discrepancies)  # No discrepancies
```

# Calculate temperature difference

Calculate temperature difference (actual value, not absolute!) between experienced temperature (temp) and mid-parental temperature.

```{r}
data_transplants<-data_transplants%>%
  mutate(mean_temp_parents=(temp_mother+temp_father)/2,
         temp_diff=temp-mean_temp_parents)
hist(data_transplants$mean_temp_parents)
hist(data_transplants$temp_diff)
```

When the temperature difference is positive --> planted in warmer places than their origin
When the temperature difference is negative --> planted in colder places than their origin

# Correct errors in fruits - seeds

Two cases with no fruits but seeds (n_tot_seed >0)! 

```{r}
data_transplants%>%
  filter(n_tot_seed>0&n_tot_fr==0)
```

Modify n_tot_seed so that it is zero when n_tot_fr is zero.

```{r}
data_transplants<-data_transplants%>%
  mutate(n_tot_seed=ifelse(n_tot_fr==0,0,n_tot_seed))
```

Cases with fruits and 0 seeds:

```{r}
nrow(data_transplants%>%filter(n_tot_seed==0&n_tot_fr>0))
```

Assign NA:

```{r}
data_transplants<-data_transplants%>%
  mutate(n_tot_seed=ifelse(n_tot_seed==0&n_tot_fr>0,NA,n_tot_seed))
```

Cases with fruits and NA seeds:

```{r}
nrow(data_transplants%>%filter(is.na(n_tot_seed==0)&n_tot_fr>0))
```

I will assign n_tot_seed as the mean of n_tot_seed of the plot for that particular year in those cases.

Calculate mean of n_tot_seed per year and plot (using only plants that produced at least one fruit):

```{r}
n_tot_seed_mean_year_plot<-data_transplants%>%filter(n_tot_fr>0)%>%
  group_by(year,plot)%>%summarise(mean=mean(n_tot_seed,na.rm=T))
```

Replace NA values with mean values:

```{r}
data_transplants <- data_transplants %>%
  left_join(n_tot_seed_mean_year_plot, by = c("year", "plot")) %>%
  mutate(n_tot_seed = ifelse(is.na(n_tot_seed) & n_tot_fr > 0, mean, n_tot_seed)) %>%
  dplyr::select(-mean)  # Remove the 'mean' column after replacement
```

# Fitness comps, both vars (Table 1, Fig 1)

I include pollination type and the quadratic effect of temperature difference in all models.

## Survival

```{r}
data_transplants_22<-data_transplants%>%
  filter(year==2022)%>%
  mutate(surv=ifelse(is.na(F_NF_A),0,1))
not_surv_22<-data_transplants_22%>%filter(surv==0)
data_transplants_23<-data_transplants%>%
  filter(year==2023)%>%
  mutate(surv=ifelse(is.na(F_NF_A),0,1))%>%
  # If the plant did not survive in 2022, set NA for surv in 2023
  mutate(surv=ifelse(unique_id %in% not_surv_22$unique_id,NA,surv))%>%
  # Correct errors: these below are in not_surv_22,
  # but actually there is data for 2023!
  # So surv for 2023 should be 1
  mutate(surv=ifelse(unique_id %in% c("Hot_10_a","Hot_30_a","Hot_45_a",
                                      "Hot_9_d","Hot_88_a","Hot_125_b"),1,surv))
# And surv for 2022 should also be 1
data_transplants_22<-data_transplants_22%>%
  mutate(surv=ifelse(unique_id %in% c("Hot_10_a","Hot_30_a","Hot_45_a",
                                      "Hot_9_d","Hot_88_a","Hot_125_b"),1,surv))
data_transplants<-rbind(data_transplants_22,data_transplants_23)
with(subset(data_transplants,year==2022),table(surv))
with(subset(data_transplants,year==2023),table(surv))
```

Of 858 plants planted in 2021, 814 survived in 2022. Of the 814 that survived in 2022, 530 survived in 2023.

```{r}
mod_surv_22<-glmmTMB(surv~temp+temp_diff+ I(temp_diff^2)+poll_type+(1|plot),
                     data_transplants%>%filter(year==2022),family="binomial")
mod_surv_23<-glmmTMB(surv~temp+temp_diff+I(temp_diff^2)+poll_type+(1|plot),
                     data_transplants%>%filter(year==2023),family="binomial")
summary(mod_surv_22)
summary(mod_surv_23)
```

### Diagnostics

```{r}
plot(simulateResiduals(mod_surv_22))
plot(simulateResiduals(mod_surv_23))
plot(check_collinearity(mod_surv_22))
plot(check_collinearity(mod_surv_23))
```

## Probability of flowering

```{r}
data_transplants<-data_transplants%>%
  mutate(flowering=ifelse(F_NF_A=="F",1,ifelse(F_NF_A=="NF",0,NA)))
data_transplants<-data_transplants%>%
  mutate(flowering=ifelse(surv==1&is.na(F_NF_A),0,flowering))
with(data_transplants,table(year,flowering))
```
 
Of 814 plants that survived in 2022, 603 flowered and 210 did not flower (and 1 was aborted, not used here). 4 were grazed before they peaked, they are included as flowering here.
Of 530 plants that survived in 2023, 411 flowered and 119 did not flower.

```{r}
mod_flowering_22<-glmmTMB(flowering~temp+temp_diff+I(temp_diff^2)+poll_type+
                            (1|plot),
                          data_transplants%>%filter(year==2022&surv==1),
                          family="binomial")
mod_flowering_23<-glmmTMB(flowering~temp+temp_diff+I(temp_diff^2)+poll_type+
                            (1|plot),
                          data_transplants%>%filter(year==2023&surv==1),
                          family="binomial")
summary(mod_flowering_22)
summary(mod_flowering_23)
```
 
### Diagnostics

```{r}
plot(simulateResiduals(mod_flowering_22))
plot(simulateResiduals(mod_flowering_23))
plot(check_collinearity(mod_flowering_22))
plot(check_collinearity(mod_flowering_23))
```

## Number of flowers

Correct tot_fl_bodies to tot_fl_bodies_corr, to be at least as high as n_tot_fr.

```{r}
nrow(data_transplants%>%filter(n_tot_fr>tot_fl_bodies))
```

221 cases where total number of fruits is larger than total number of flowering bodies.

```{r}
data_transplants<-data_transplants%>%
  mutate(tot_fl_bodies_corr=ifelse(n_tot_fr>tot_fl_bodies,n_tot_fr,
                                   tot_fl_bodies),
         diff_fr_fl=n_tot_fr-tot_fl_bodies)
```

Histogram of differences:

```{r}
hist(data_transplants$diff_fr_fl)
```

Negative differences: more flowers than fruits - OK
Positive differences: more fruits than flowers - not OK --> corrected

```{r}
hist(subset(data_transplants,year==2022)$tot_fl_bodies_corr)
hist(subset(data_transplants,year==2022&flowering==1)$tot_fl_bodies_corr)
hist(subset(data_transplants,year==2023)$tot_fl_bodies_corr)
hist(subset(data_transplants,year==2023&flowering==1)$tot_fl_bodies_corr)
```

Models for number of flowers in those that flowered (flowering=1).

#### Without median_h

```{r}
mod_nfl_22<-glmmTMB(tot_fl_bodies_corr~temp+temp_diff+
                            I(temp_diff^2)+poll_type+(1|plot),
                          data_transplants%>%filter(year==2022&flowering==1),
                          family="nbinom2")
mod_nfl_23<-glmmTMB(tot_fl_bodies_corr~temp+temp_diff+I(temp_diff^2)+ 
                            poll_type+(1|plot),
                          data_transplants%>%filter(year==2023&flowering==1),
                          family="nbinom2")
summary(mod_nfl_22)
summary(mod_nfl_23)
``` 

##### Diagnostics

```{r}
plot(simulateResiduals(mod_nfl_22))
testDispersion(simulateResiduals(mod_nfl_22),alternative="greater")
plot(simulateResiduals(mod_nfl_23))
testDispersion(simulateResiduals(mod_nfl_23),alternative="greater")
plot(check_collinearity(mod_nfl_22))
plot(check_collinearity(mod_nfl_23))
```

#### With median_h

```{r}
mod_nfl_22_mh<-glmmTMB(tot_fl_bodies_corr~temp+temp_diff+I(temp_diff^2)+ 
                               median_h+
                               poll_type+(1|plot),
                             data_transplants%>%filter(year==2022&flowering==1),
                             family="nbinom2")
mod_nfl_23_mh<-glmmTMB(tot_fl_bodies_corr~temp+temp_diff+I(temp_diff^2)+ 
                               median_h+
                               poll_type+(1|plot),
                             data_transplants%>%filter(year==2023&flowering==1),
                             family="nbinom2")
summary(mod_nfl_22_mh)
summary(mod_nfl_23_mh)
``` 

##### Diagnostics

```{r}
plot(simulateResiduals(mod_nfl_22_mh))
testDispersion(simulateResiduals(mod_nfl_22_mh),alternative="greater")
plot(simulateResiduals(mod_nfl_23_mh))
testDispersion(simulateResiduals(mod_nfl_23_mh),alternative="greater")
plot(check_collinearity(mod_nfl_22_mh))
plot(check_collinearity(mod_nfl_23_mh))
```

## Probability of fruiting

```{r}
data_transplants<-data_transplants%>%
  mutate(fruiting=ifelse(surv==0|is.na(surv),NA, 
                         # NA for plants that did not survive
                         ifelse(F_NF_A=="A"|F_NF_A=="NF",NA, 
                                # NA for plants that aborted or did not flower
                                ifelse(is.na(n_tot_fr),NA,
                                       # NA for plants where n_tot_fr is NA
                                       ifelse(flowering==0,NA,
                                              # NA for plants that did not flower
                                              ifelse(n_tot_fr>0,1,0)))))) 
# 1/0 for plants that produced / did not produce fruits
table(subset(data_transplants,year==2022)$fruiting)
table(subset(data_transplants,year==2023)$fruiting)
hist(data_transplants$fruiting)
```

Of 603 plants that flowered in 2022, 582 produced fruits and 21 did not produce fruits.
Of 411 plants that flowered in 2022, 398 produced fruits and 13 did not produce fruits.

```{r}
mod_fruiting_22<-glmmTMB(fruiting~temp+temp_diff+I(temp_diff^2)+poll_type+ 
                           (1|plot),
                         data_transplants%>%filter(year==2022&flowering==1),
                         family="binomial")
mod_fruiting_23<-glmmTMB(fruiting~temp+temp_diff+I(temp_diff^2)+poll_type+
                           (1|plot),
                         data_transplants%>%filter(year==2023&flowering==1),
                         family="binomial")
summary(mod_fruiting_22)
summary(mod_fruiting_23)
```

### Diagnostics

```{r}
plot(simulateResiduals(mod_fruiting_22))
plot(simulateResiduals(mod_fruiting_23))
plot(check_collinearity(mod_fruiting_22))
plot(check_collinearity(mod_fruiting_23))
```

## Number of fruits

Models for number of fruits in those that flowered (flowering=1).

```{r}
mod_nfr_22<-glmmTMB(n_tot_fr~temp+temp_diff+I(temp_diff^2)+
                            poll_type+(1|plot),
                          data_transplants%>%filter(year==2022&flowering==1),
                          family="nbinom2")
mod_nfr_23<-glmmTMB(n_tot_fr~temp+temp_diff+I(temp_diff^2)+ 
                            poll_type+(1|plot),
                          data_transplants%>%filter(year==2023&flowering==1),
                          family="nbinom2")
summary(mod_nfr_22)
summary(mod_nfr_23)
``` 

### Diagnostics

```{r}
plot(simulateResiduals(mod_nfr_22))
testDispersion(simulateResiduals(mod_nfr_22),alternative="greater")
plot(simulateResiduals(mod_nfr_23))
testDispersion(simulateResiduals(mod_nfr_23),alternative="greater")
plot(check_collinearity(mod_nfr_22))
plot(check_collinearity(mod_nfr_23))
```

## Number of seeds

Including only plants that produced fruits.

```{r}
hist((data_transplants%>%filter(fruiting==1))$n_tot_seed)
```

```{r}
mod_nseed_22<-glmmTMB(n_tot_seed~temp+temp_diff+I(temp_diff^2)+ 
                            poll_type+(1|plot),
                          data_transplants%>%filter(year==2022&fruiting==1),
                          family="nbinom2")
mod_nseed_23<-glmmTMB(n_tot_seed~temp+temp_diff+I(temp_diff^2)+ 
                            poll_type+(1|plot),
                          data_transplants%>%filter(year==2023&fruiting==1),
                          family="nbinom2")
summary(mod_nseed_22)
summary(mod_nseed_23)
``` 

### Diagnostics

```{r}
plot(simulateResiduals(mod_nseed_22))
plot(simulateResiduals(mod_nseed_23))
testDispersion(simulateResiduals(mod_nseed_22),alternative="greater")
testDispersion(simulateResiduals(mod_nseed_23),alternative="greater")
plot(check_collinearity(mod_nseed_22))
plot(check_collinearity(mod_nseed_23))
```
## Table 1

```{r}
tab_model(mod_surv_22,mod_flowering_22,mod_nfl_22,
          mod_nfl_22_mh,mod_fruiting_22,mod_nfr_22,
          mod_nseed_22,
          transform=NULL,show.intercept=F,show.ci=F,show.se=F,show.stat=F,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("Survival","Prob. of flowering",
                      "N flowers (1)","N flowers (2)","Prob. of fruiting",
                      "N fruits","N seeds"),
          file="output/tables/summary_2022.doc")
tab_model(mod_surv_23,mod_flowering_23,
          mod_nfl_23,mod_nfl_23_mh,mod_fruiting_23,mod_nfr_23,
          mod_nseed_23,
          transform=NULL,show.intercept=F,show.ci=F,show.se=F,show.stat=F,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("Survival","Prob. of flowering",
                      "N flowers (1)","N flowers (2)","Prob. of fruiting",
                      "N fruits","N seeds"),
          file="output/tables/summary_2023.doc")
```

## Figure 1

```{r}
fig1_1<-
  # surv temp
  ggplot()+
  geom_jitter(data=data_transplants,aes(x=temp,y=surv,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_surv_22,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_surv_22,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_surv_23,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_surv_23,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5)+
  my_theme()+xlab(NULL)+ylab(NULL)+
  scale_x_continuous(limits=c(0,50),breaks=seq(0,50,10))
fig1_2<-
  # surv temp_diff
  ggplot()+
  geom_jitter(data=data_transplants,aes(x=temp_diff,y=surv,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_surv_22,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_surv_22,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5,
            linetype="dashed")+
        geom_ribbon(data=ggpredict(mod_surv_23,terms=c("temp_diff[all]")),
                    aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
                    fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_surv_23,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5)+
  my_theme()+xlab(NULL)+ylab(NULL)
fig1_3<-
  # flowering temp
  ggplot()+
  geom_jitter(data=data_transplants%>%filter(surv==1),
             aes(x=temp,y=flowering,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_flowering_22,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_flowering_22,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_flowering_23,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_flowering_23,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5)+
  my_theme()+xlab(NULL)+ylab(NULL)+
  scale_x_continuous(limits=c(0,50),breaks=seq(0,50,10))
fig1_4<-
  # flowering temp_diff
  ggplot()+
  geom_jitter(data=data_transplants%>%filter(surv==1),
             aes(x=temp_diff,y=flowering,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_flowering_22,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_flowering_22,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5,
            linetype="dashed")+
  geom_ribbon(data=ggpredict(mod_flowering_23,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_flowering_23,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5,
            linetype="dashed")+
  my_theme()+xlab(NULL)+ylab(NULL)
fig1_5<-
  # fruiting temp
  ggplot()+
  geom_jitter(data=data_transplants%>%filter(flowering==1),
             aes(x=temp,y=fruiting,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_fruiting_22,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_fruiting_22,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_fruiting_23,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_fruiting_23,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5,linetype="dashed")+
  my_theme()+xlab(NULL)+ylab(NULL)+
  scale_x_continuous(limits=c(0,50),breaks=seq(0,50,10))
fig1_6<-
  # fruiting temp_diff
  ggplot()+
  geom_jitter(data=data_transplants%>%filter(flowering==1),
             aes(x=temp_diff,y=fruiting,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_fruiting_22,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_fruiting_22,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_fruiting_23,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_fruiting_23,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5,linetype="dashed")+
  my_theme()+xlab(NULL)+ylab(NULL)
fig1_7<-
  # N seeds temp
  ggplot()+
  geom_point(data=data_transplants%>%filter(fruiting==1),
             aes(x=temp,y=n_tot_seed,color=year),
              size=1,alpha=0.2)+
  geom_ribbon(data=ggpredict(mod_nseed_22,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_nseed_22,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_nseed_23,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_nseed_23,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5,linetype="dashed")+
  my_theme()+xlab(NULL)+ylab(NULL)+
  scale_x_continuous(limits=c(0,50),breaks=seq(0,50,10))
fig1_8<-  
  # N seeds temp_diff
  ggplot()+
  geom_point(data=data_transplants%>%filter(fruiting==1),
             aes(x=temp_diff,y=n_tot_seed,color=year),
              size=1,alpha=0.2)+
  geom_ribbon(data=ggpredict(mod_nseed_22,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_nseed_22,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_nseed_23,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_nseed_23,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5,linetype="dashed")+
  my_theme()+xlab(NULL)+ ylab(NULL)
```

```{r}
fig1<-grid.arrange(
  arrangeGrob(
    arrangeGrob(fig1_1,fig1_2,ncol=2,
                left=textGrob("Survival",rot=90,
                              gp=gpar(fontsize=12,fontfamily="serif"),
                              just="center")),
    arrangeGrob(fig1_3,fig1_4,ncol=2,
                left=textGrob("      Flowering incidence",rot=90,
                              gp=gpar(fontsize=12,fontfamily="serif"),
                              just="center")),
    arrangeGrob(fig1_5,fig1_6,ncol=2,
                left=textGrob("      Fruiting incidence",rot=90,
                              gp=gpar(fontsize=12,fontfamily="serif"),
                              just="center")),
    arrangeGrob(fig1_7,fig1_8,ncol=2,
                left=textGrob("      Total number of seeds",rot=90,
                              gp=gpar(fontsize=12,fontfamily="serif"),
                              just="center")),
    nrow = 4,
    bottom=arrangeGrob(
      textGrob("                    Temperature at planting site (ºC)",
               gp=gpar(fontsize=12,fontfamily="serif"),just="center"),
      textGrob("              Temperature difference (ºC)",
               gp=gpar(fontsize=12,fontfamily="serif"),just="center"),
      ncol = 2,
      heights = unit.c(unit(1, "lines"), unit(0.5, "lines")))
    )
  )
ggsave(file="output/figures/fig1.tiff", plot=fig1,width=6,height=8,dpi=300,
       compression="lzw")
```

## Figures pollination - needed?

```{r}
ggplot()+
  geom_point(data=ggpredict(mod_surv_23,terms=c("poll_type[all]")),
             aes(x=x,y=predicted),size=3)+
  geom_errorbar(data=ggpredict(mod_surv_23,terms=c("poll_type[all]")),
                aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),width=0.1)+
  my_theme()+xlab("Pollination type")+ylab("Probability of survival")+
  ggtitle("2023")

grid.arrange(
  ggplot()+
    #geom_jitter(data=data_transplants%>%filter(year==2022),
    #            aes(x=poll_type,y=flowering),width=0.05,alpha=0.1)+
    geom_point(data=ggpredict(mod_flowering_22,terms=c("poll_type[all]")),
               aes(x=x,y=predicted),size=3)+
    geom_errorbar(data=ggpredict(mod_flowering_22,terms=c("poll_type[all]")),
                  aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
                  width=0.1)+
    my_theme()+xlab("Pollination type")+ylab("Probability of flowering")+
    ggtitle("2022"),
  ggplot()+
    #geom_jitter(data=data_transplants%>%filter(year==2023),
    #            aes(x=poll_type,y=flowering),width=0.05,alpha=0.1)+
    geom_point(data=ggpredict(mod_flowering_23,terms=c("poll_type[all]")),
               aes(x=x,y=predicted),size=3)+
    geom_errorbar(data=ggpredict(mod_flowering_23,terms=c("poll_type[all]")),
                  aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
                  width=0.1)+
    my_theme()+xlab("Pollination type")+ylab("Probability of flowering")+
    ggtitle("2023"),
ncol=2)

ggplot()+
  geom_point(data=ggpredict(mod_nseed_22,terms=c("poll_type[all]")),
             aes(x=x,y=predicted),size=3)+
  geom_errorbar(data=ggpredict(mod_nseed_22,terms=c("poll_type[all]")),
                aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),width=0.1)+ 
  my_theme()+xlab("Pollination type")+ylab("Total number of seeds")+
  ggtitle("2022")
```

# Appendix 2: Distributions fitness components

```{r}
App2<-grid.arrange(
  ggplot(data_transplants%>%filter(!is.na(surv)),
         aes(x=factor(surv)))+
    geom_bar(color="darkgrey",fill="grey",alpha=0.6)+facet_wrap(~year)+
    my_theme()+xlab("Survival")+ylab(NULL),
  ggplot(data_transplants%>%filter(surv==1&!is.na(flowering)),
         aes(x=factor(flowering)))+
    geom_bar(color="darkgrey",fill="grey",alpha=0.6)+facet_wrap(~year)+
    my_theme()+xlab("Flowering incidence")+ylab(NULL),
  ggplot(data_transplants%>%filter(flowering==1),
         aes(x=factor(fruiting)))+
    geom_bar(color="darkgrey",fill="grey",alpha=0.6)+facet_wrap(~year)+
    my_theme()+xlab("Fruiting incidence")+ylab(NULL),
  ggplot(data_transplants%>%filter(fruiting==1),
         aes(x=n_tot_seed))+
    geom_histogram(bins=20,color="darkgrey",fill="grey",alpha=0.6)+
    facet_wrap(~year)+
    my_theme()+xlab("Total number of seeds")+ylab(NULL),
  ncol=2,
  left=textGrob("Count",rot=90,
                gp=gpar(fontsize=16,fontfamily="serif"),just="center"))
ggsave(file="output/figures/App2.tiff", plot=App2,width=10,height=8,dpi=300,
       compression="lzw")
```


# Appendix 3: Differences in fitness components among years

```{r warning=FALSE}
mod_surv_years<-glm(surv~year,data_transplants,family="binomial")
mod_flowering_years<-glm(flowering~year,data_transplants%>%filter(surv==1),
                         family="binomial")
mod_fruiting_years<-glm(fruiting~year,data_transplants%>%filter(flowering==1),
                        family="binomial")
mod_nseed_years<-glm.nb(n_tot_seed~year,data_transplants%>%filter(fruiting==1))
```

```{r}
summary(mod_surv_years) #*
summary(mod_flowering_years)
summary(mod_fruiting_years)
summary(mod_nseed_years) #*
```

```{r}
plot(ggpredict(mod_surv_years)) #*
plot(ggpredict(mod_flowering_years))
plot(ggpredict(mod_fruiting_years))
plot(ggpredict(mod_nseed_years)) #*
```

```{r}
App3<-grid.arrange(
  ggplot(data_transplants%>%group_by(year)%>%
           summarize(mean_value=mean(surv,na.rm=T),
                     se_value=sd(surv,na.rm=T)/sqrt(n())),
         aes(x=factor(year),y= mean_value))+geom_point(size=4)+
    geom_errorbar(aes(ymin=mean_value-se_value,ymax=mean_value+se_value),
                  width=0.2)+
    xlab("Year")+ylab("Survival")+my_theme(),
  ggplot(data_transplants%>%filter(surv==1)%>%group_by(year)%>%
           summarize(mean_value=mean(flowering,na.rm=T),
                     se_value=sd(flowering,na.rm=T)/sqrt(n())),
         aes(x=factor(year),y= mean_value))+geom_point(size=4)+
    geom_errorbar(aes(ymin=mean_value-se_value,ymax=mean_value+se_value),
                  width=0.2)+
    xlab("Year")+ylab("Flowering incidence")+my_theme(),
  ggplot(data_transplants%>%filter(flowering==1)%>%group_by(year)%>%
           summarize(mean_value=mean(fruiting,na.rm=T),
                     se_value=sd(fruiting,na.rm=T)/sqrt(n())),
         aes(x=factor(year),y= mean_value))+geom_point(size=4)+
    geom_errorbar(aes(ymin=mean_value-se_value,ymax=mean_value+se_value),
                  width=0.2)+
    xlab("Year")+ylab("Fruiting incidence")+my_theme(),
  ggplot(data_transplants%>%filter(fruiting==1)%>%group_by(year)%>%
           summarize(mean_value=mean(n_tot_seed,na.rm=T),
                     se_value=sd(n_tot_seed,na.rm=T)/sqrt(n())),
         aes(x=factor(year),y= mean_value))+geom_point(size=4)+
    geom_errorbar(aes(ymin=mean_value-se_value,ymax=mean_value+se_value),
                  width=0.2)+
    xlab("Year")+ylab("Total number of seeds")+my_theme(),
  ncol=2)
ggsave(file="output/figures/App3.tiff", plot=App3,width=10,height=8,dpi=300,
       compression="lzw")
```


# Path model n seeds

Path models with fitness component number of seeds, to see the direct effects of temp and temp_diff vs. the indirect effects via FFD. Path models are not possible to build with Aster models because FFD is NA when surv = 0 and / or flowering = 0.

I use GLMMs with plot as random factor.

## Fit component models

```{r}
data_path_22<-data_transplants%>%filter(year==2022&fruiting==1)%>%
  mutate(temp_diff_sq=I(temp_diff^2))
data_path_23<-data_transplants%>%filter(year==2023&fruiting==1)%>%
  mutate(temp_diff_sq=I(temp_diff^2))
```

```{r}
mod_nseed_22_path1<-glmmTMB(FFD_corr~temp+temp_diff+temp_diff_sq+poll_type+
                              (1|plot),data_path_22,family="gaussian")
mod_nseed_22_path2<-glmmTMB(n_tot_seed~temp+temp_diff+temp_diff_sq+poll_type+
                              FFD_corr+(1|plot),data_path_22,family="nbinom2")
mod_nseed_23_path1<-glmmTMB(FFD_corr~temp+temp_diff+temp_diff_sq+poll_type+
                              (1|plot),data_path_23,family="gaussian")
mod_nseed_23_path2<-glmmTMB(n_tot_seed~temp+temp_diff+temp_diff_sq+poll_type+
                              FFD_corr+(1|plot),data_path_23,family="nbinom2")
```

## Diagnostics

```{r}
plot(simulateResiduals(mod_nseed_22_path1)) # KS* but how to improve?
plot(simulateResiduals(mod_nseed_22_path2))
plot(check_collinearity(mod_nseed_22_path1))
plot(check_collinearity(mod_nseed_22_path2))
plot(simulateResiduals(mod_nseed_23_path1)) # KS* but how to improve?
plot(simulateResiduals(mod_nseed_23_path2))
plot(check_collinearity(mod_nseed_23_path1))
plot(check_collinearity(mod_nseed_23_path2))
```

## Fit path models

```{r}
path_nseed_22<-psem(mod_nseed_22_path1,mod_nseed_22_path2)
path_nseed_23<-psem(mod_nseed_23_path1,mod_nseed_23_path2)
```

```{r}
coefs(path_nseed_22)
coefs(path_nseed_23)
```

# Path model n seeds in flowering ids

Path models with fitness component fruiting x fitness component number of seeds, to see the direct effects of temp and temp_diff vs. the indirect effects via FFD. Path models are not possible to build with Aster models because FFD is NA when surv = 0 and / or flowering = 0.

I use GLMMs with plot as random factor.

## Fit component models

```{r}
data_path_22_in_fl<-data_transplants%>%filter(year==2022&flowering==1)%>%
  mutate(temp_diff_sq=I(temp_diff^2))
data_path_23_in_fl<-data_transplants%>%filter(year==2023&flowering==1)%>%
  mutate(temp_diff_sq=I(temp_diff^2))
```

```{r}
mod_nseed_22_in_fl_path1<-glmmTMB(FFD_corr~temp+temp_diff+temp_diff_sq+
                                    poll_type+(1|plot),
                                  data_path_22_in_fl,family="gaussian")
mod_nseed_22_in_fl_path2<-glmmTMB(n_tot_seed~temp+temp_diff+temp_diff_sq+
                                    poll_type+FFD_corr+(1|plot),
                                  data_path_22_in_fl,family="nbinom2")
mod_nseed_23_in_fl_path1<-glmmTMB(FFD_corr~temp+temp_diff+temp_diff_sq+
                                    poll_type+(1|plot),
                                  data_path_23_in_fl,family="gaussian")
mod_nseed_23_in_fl_path2<-glmmTMB(n_tot_seed~temp+temp_diff+temp_diff_sq+
                                    poll_type+FFD_corr+(1|plot),
                                  data_path_23_in_fl,family="nbinom2")
```

## Diagnostics

```{r}
plot(simulateResiduals(mod_nseed_22_in_fl_path1)) # KS* but how to improve?
plot(simulateResiduals(mod_nseed_22_in_fl_path2))
plot(check_collinearity(mod_nseed_22_in_fl_path1))
plot(check_collinearity(mod_nseed_22_in_fl_path2))
plot(simulateResiduals(mod_nseed_23_in_fl_path1)) # KS* but how to improve?
plot(simulateResiduals(mod_nseed_23_in_fl_path2))
plot(check_collinearity(mod_nseed_23_in_fl_path1))
plot(check_collinearity(mod_nseed_23_in_fl_path2))
```

## Fit path models

```{r}
path_nseed_22_in_fl<-psem(mod_nseed_22_in_fl_path1,mod_nseed_22_in_fl_path2,
                          temp %~~% temp_diff)
path_nseed_23_in_fl<-psem(mod_nseed_23_in_fl_path1,mod_nseed_23_in_fl_path2)
```

```{r}
coefs(path_nseed_22_in_fl)
coefs(path_nseed_23_in_fl)
```

# Correlation temp - temp_diff

```{r}
with(data_transplants,cor(temp,temp_diff,use="complete.obs"))
with(data_transplants%>%filter(year==2022),cor(temp,temp_diff,
                                              use="complete.obs"))
with(data_transplants%>%filter(year==2023),cor(temp,temp_diff,
                                               use="complete.obs"))
```

```{r}
ggplot(data_transplants,aes(x=temp,y=temp_diff))+geom_point()+
  geom_smooth(method="lm")+facet_wrap(~year)
ggplot(data_transplants,aes(x=temp,y=abs(temp_diff)))+geom_point()+
  geom_smooth(method="lm")+facet_wrap(~year)+
  geom_smooth(method="lm",formula=y ~ x + I(x^2),color="red")
```

# Fitness comps, single vars (Apps 1-2)

## Survival

```{r}
mod_surv_22_temp<-glmmTMB(surv~temp+poll_type+(1|plot),
                     data_transplants%>%filter(year==2022),family="binomial")
mod_surv_23_temp<-glmmTMB(surv~temp+poll_type+(1|plot),
                     data_transplants%>%filter(year==2023),family="binomial")
summary(mod_surv_22_temp)
summary(mod_surv_23_temp)
```

```{r}
mod_surv_22_temp_diff<-glmmTMB(surv~temp_diff+I(temp_diff^2)+poll_type+(1|plot),
                     data_transplants%>%filter(year==2022),family="binomial")
mod_surv_23_temp_diff<-glmmTMB(surv~temp_diff+I(temp_diff^2)+poll_type+(1|plot),
                     data_transplants%>%filter(year==2023),family="binomial")
summary(mod_surv_22_temp_diff)
summary(mod_surv_23_temp_diff)
```

## Probability of flowering

```{r}
mod_flowering_22_temp<-glmmTMB(flowering~temp+poll_type+
                            (1|plot),data_transplants%>%filter(year==2022),
                          family="binomial")
mod_flowering_23_temp<-glmmTMB(flowering~temp+poll_type+
                            (1|plot),data_transplants%>%filter(year==2023),
                          family="binomial")
summary(mod_flowering_22_temp)
summary(mod_flowering_23_temp)
```

```{r}
mod_flowering_22_temp_diff<-glmmTMB(flowering~temp_diff+I(temp_diff^2)+
                            poll_type+(1|plot),
                            data_transplants%>%filter(year==2022),
                          family="binomial")
mod_flowering_23_temp_diff<-glmmTMB(flowering~temp_diff+I(temp_diff^2)+
                            poll_type+(1|plot),
                            data_transplants%>%filter(year==2023),
                          family="binomial")
summary(mod_flowering_22_temp_diff)
summary(mod_flowering_23_temp_diff)
```

## Probability of fruiting

```{r}
mod_fruiting_22_temp<-glmmTMB(fruiting~temp+poll_type+
                           (1|plot),data_transplants%>%filter(year==2022),
                         family="binomial")
mod_fruiting_23_temp<-glmmTMB(fruiting~temp+poll_type+
                           (1|plot),data_transplants%>%filter(year==2023),
                         family="binomial")
summary(mod_fruiting_22_temp)
summary(mod_fruiting_23_temp)
```

```{r}
mod_fruiting_22_temp_diff<-glmmTMB(fruiting~temp_diff+I(temp_diff^2)+ 
                           poll_type+(1|plot),
                           data_transplants%>%filter(year==2022),
                         family="binomial")
mod_fruiting_23_temp_diff<-glmmTMB(fruiting~temp_diff+I(temp_diff^2)+
                           poll_type+(1|plot),
                           data_transplants%>%filter(year==2023),
                         family="binomial")
summary(mod_fruiting_22_temp_diff)
summary(mod_fruiting_23_temp_diff)
```

## Number of seeds

```{r}
mod_nseed_22_temp<-glmmTMB(n_tot_seed~temp+poll_type+(1|plot),
                          data_transplants%>%filter(year==2022&fruiting==1),
                          family="nbinom2")
mod_nseed_23_temp<-glmmTMB(n_tot_seed~temp+poll_type+(1|plot),
                          data_transplants%>%filter(year==2023&fruiting==1),
                          family="nbinom2")
summary(mod_nseed_22_temp)
summary(mod_nseed_23_temp)
``` 

```{r}
mod_nseed_22_temp_diff<-glmmTMB(n_tot_seed~temp_diff+I(temp_diff^2)+poll_type+ 
                            (1|plot),
                          data_transplants%>%filter(year==2022&fruiting==1),
                          family="nbinom2")
mod_nseed_23_temp_diff<-glmmTMB(n_tot_seed~temp_diff+I(temp_diff^2)+poll_type+ 
                            (1|plot),
                          data_transplants%>%filter(year==2023&fruiting==1),
                          family="nbinom2")
summary(mod_nseed_22_temp_diff)
summary(mod_nseed_23_temp_diff)
``` 

## Appendix 4

```{r}
tab_model(mod_surv_22_temp,mod_flowering_22_temp,mod_fruiting_22_temp,
          mod_nseed_22_temp,
          transform=NULL,show.intercept=F,show.ci=F,show.se=F,show.stat=F,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("Survival","Prob. of flowering","Prob. of fruiting",
                      "N seeds"),
          file="output/tables/summary_2022_temp.doc")
```

```{r}
tab_model(mod_surv_22_temp_diff,mod_flowering_22_temp_diff,
          mod_fruiting_22_temp_diff,mod_nseed_22_temp_diff,
          transform=NULL,show.intercept=F,show.ci=F,show.se=F,show.stat=F,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("Survival","Prob. of flowering","Prob. of fruiting",
                      "N seeds"),
          file="output/tables/summary_2022_temp_diff.doc")
```

```{r}
tab_model(mod_surv_23_temp,mod_flowering_23_temp,mod_fruiting_23_temp,
          mod_nseed_23_temp,
          transform=NULL,show.intercept=F,show.ci=F,show.se=F,show.stat=F,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("Survival","Prob. of flowering","Prob. of fruiting",
                      "N seeds"),
          file="output/tables/summary_2023_temp.doc")
```

```{r}
tab_model(mod_surv_23_temp_diff,mod_flowering_23_temp_diff,
          mod_fruiting_23_temp_diff,mod_nseed_23_temp_diff,
          transform=NULL,show.intercept=F,show.ci=F,show.se=F,show.stat=F,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("Survival","Prob. of flowering","Prob. of fruiting",
                      "N seeds"),
          file="output/tables/summary_2023_temp_diff.doc")
```

## Appendix 5

```{r}
App5_1<-ggplot()+
  geom_jitter(data=data_transplants,aes(x=temp,y=surv,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_surv_22_temp,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_surv_22_temp,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_surv_23_temp,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_surv_23_temp,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5)+
  my_theme()+xlab(NULL)+ylab(NULL)+
  scale_x_continuous(limits=c(0,50),breaks=seq(0,50,10))
App5_2<-ggplot()+
  geom_jitter(data=data_transplants,aes(x=temp_diff,y=surv,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_surv_22_temp_diff,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_surv_22_temp_diff,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_surv_23_temp_diff,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_surv_23_temp_diff,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5)+
  my_theme()+xlab(NULL)+ylab(NULL)
App5_3<-ggplot()+
  geom_jitter(data=data_transplants,aes(x=temp,y=flowering,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_flowering_22_temp,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_flowering_22_temp,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_flowering_23_temp,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_flowering_23_temp,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5)+
  my_theme()+xlab(NULL)+ylab(NULL)+
  scale_x_continuous(limits=c(0,50),breaks=seq(0,50,10))
App5_4<-ggplot()+
  geom_jitter(data=data_transplants,aes(x=temp_diff,y=flowering,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_flowering_22_temp_diff,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_flowering_22_temp_diff,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_flowering_23_temp_diff,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_flowering_23_temp_diff,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5,linetype="dashed")+
  my_theme()+xlab(NULL)+ylab(NULL)
App5_5<-ggplot()+
  geom_jitter(data=data_transplants,aes(x=temp,y=fruiting,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_fruiting_22_temp,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_fruiting_22_temp,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5,linetype="dashed")+
  geom_ribbon(data=ggpredict(mod_fruiting_23_temp,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_fruiting_23_temp,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5,linetype="dashed")+
  my_theme()+xlab(NULL)+ylab(NULL)+
  scale_x_continuous(limits=c(0,50),breaks=seq(0,50,10))
App5_6<-ggplot()+
  geom_jitter(data=data_transplants,aes(x=temp_diff,y=fruiting,color=year),
              size=1,alpha=0.2,width=0,height=0.04)+
  geom_ribbon(data=ggpredict(mod_fruiting_22_temp_diff,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_fruiting_22_temp_diff,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5,linetype="dashed")+
  geom_ribbon(data=ggpredict(mod_fruiting_23_temp_diff,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_fruiting_23_temp_diff,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5,linetype="dashed")+
  my_theme()+xlab(NULL)+ylab(NULL)
App5_7<-ggplot()+
  geom_point(data=data_transplants%>%filter(fruiting==1),
             aes(x=temp,y=n_tot_seed,color=year),size=1,alpha=0.2)+
  geom_ribbon(data=ggpredict(mod_nseed_22_temp,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_nseed_22_temp,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5)+
  geom_ribbon(data=ggpredict(mod_nseed_23_temp,terms=c("temp[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_nseed_23_temp,terms=c("temp[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5,linetype="dashed")+
  my_theme()+xlab(NULL)+ylab(NULL)+
  scale_x_continuous(limits=c(0,50),breaks=seq(0,50,10))
App5_8<-ggplot()+
  geom_point(data=data_transplants%>%filter(fruiting==1),
             aes(x=temp_diff,y=n_tot_seed,color=year),size=1,alpha=0.2)+
  geom_ribbon(data=ggpredict(mod_nseed_22_temp_diff,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#F8766D",alpha=0.2)+
  geom_line(data=ggpredict(mod_nseed_22_temp_diff,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#F8766D",size=0.5,linetype="dashed")+
  geom_ribbon(data=ggpredict(mod_nseed_23_temp_diff,terms=c("temp_diff[all]")),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="#00BFC4",alpha=0.2)+
  geom_line(data=ggpredict(mod_nseed_23_temp_diff,terms=c("temp_diff[all]")),
            aes(x=x,y=predicted),color="#00BFC4",size=0.5)+
  my_theme()+xlab(NULL)+ylab(NULL)
```

```{r}
App5<-grid.arrange(
  arrangeGrob(
    arrangeGrob(App5_1,App5_2,ncol=2,
                left=textGrob("      Survival",rot=90,
                              gp=gpar(fontsize=12,fontfamily="serif"),
                              just="center")),
    arrangeGrob(App5_3,App5_4,ncol=2,
                left=textGrob("      Flowering incidence",rot=90,
                              gp=gpar(fontsize=12,fontfamily="serif"),
                              just="center")),
    arrangeGrob(App5_5,App5_6,ncol=2,
                left=textGrob("      Fruiting incidence",rot=90,
                              gp=gpar(fontsize=12,fontfamily="serif"),
                              just="center")),
    arrangeGrob(App5_7,App5_8,ncol=2,
                left=textGrob("      Total number of seeds",rot=90,
                              gp=gpar(fontsize=12,fontfamily="serif"),
                              just="center")),
    nrow = 4,
    bottom=arrangeGrob(
      textGrob("                    Temperature at planting site (ºC)",
               gp=gpar(fontsize=12,fontfamily="serif"),just="center"),
      textGrob("              Temperature difference (ºC)",
               gp=gpar(fontsize=12,fontfamily="serif"),just="center"),
      ncol = 2,
      heights = unit.c(unit(1, "lines"), unit(0.5, "lines")))
    )
  )
ggsave(file="output/figures/App5.tiff", plot=App5,width=6,height=8,dpi=300,
       compression="lzw")
```

# Aster data prep

Using multiplicative fitness components.

Survival --> Any flowers --> Any Fruits --> Seeds

Survival (did the plant survive?) is Bernouilli.

Any flowers (did the plant produce flowers?) is Bernouilli.

Any fruits (did the plant produce fruits?) is Bernoulli.

Number of seeds is zero-truncated negative binomial / zero-truncated poisson, because seed number cannot be zero if fruit production = 1.

Remove 5 plants that were grazed before they peaked in 2022, 1 plant that was aborted, and some plants where temp or temp_diff (or both) were NA:

```{r}
data_transplants_aster<-data_transplants%>%
  filter(comments_peak!="Grazed before it peaked"|is.na(comments_peak))%>%
  filter(F_NF_A!="A"|is.na(F_NF_A))%>%
  filter(!is.na(temp_diff)&!is.na(temp))
```

```{r}
range((data_transplants_aster%>%filter(n_tot_fr>0))$n_tot_seed)
```

Number of seeds (in those that produced fruits) ranges from 1 to 221.

```{r}
data_transplants_aster$nseed<-with(
  data_transplants_aster,
  ifelse(surv==0,0,
         ifelse(flowering==0,0,
                ifelse(fruiting==0,0,
                       as.integer(data_transplants_aster$n_tot_seed)))))
data_transplants_aster$fl_01<-with(data_transplants_aster,
                                   ifelse(is.na(flowering),0,flowering))
data_transplants_aster$fr_01<-with(data_transplants_aster,
                                   ifelse(is.na(fruiting),0,fruiting))
vars <- c("surv", "fl_01","fr_01", "nseed") # Our fitness variables
```

Show the graphical model:

```{r}
pred <- c(0,1,2,3) # specifies the predecessor structure of the graph

foo <- c("root", vars)
pvars <- foo[pred + 1]
bar <- cbind(pvars, vars)
colnames(bar) <- c("pred", "succ")
bar
```

## 2022

Reshape data:

```{r}
redata22 <- reshape(data.frame(data_transplants_aster%>%filter(year==2022)%>%
                      dplyr::select(unique_id,plot,poll_type,temp,temp_diff,
                                    FFD_corr,surv,fl_01,fr_01,nseed)), 
                  varying = list(vars),
                  direction = "long", 
                  timevar = "varb", # reshape to long format
                  idvar = "unique_id",
                  times = as.factor(vars), v.names = "resp")
redata22 <- data.frame(redata22, root = 1) # This adds a variable root to the df
# and makes all its values one (including for non-root nodes, but those values
# are ignored by all aster package functions).

# resp contains all of the data in the variables indicated by the string
# vars packed into a single vector
# varb indicates which original variable the corresponding element of
# resp came from
```

```{r}
names(data.frame(data_transplants_aster%>%filter(year==2022)%>%
                      dplyr::select(unique_id,plot,poll_type,temp,temp_diff,
                                    FFD_corr,surv,fl_01,fr_01,nseed)))
names(redata22)
nrow(data.frame(data_transplants_aster%>%filter(year==2022)%>%
                      dplyr::select(unique_id,plot,poll_type,temp,temp_diff,
                                    FFD_corr,surv,fl_01,fr_01,nseed)))
nrow(redata22)
nrow(data.frame(data_transplants_aster%>%filter(year==2022)%>%
                  dplyr::select(unique_id,plot,poll_type,temp,temp_diff,
                                FFD_corr,surv,fl_01,fr_01,nseed)))*length(vars)
sapply(redata22, class)
levels(redata22$varb)
length(unique(redata22$unique_id))
```

Look at Poisson distributions for fruits:

```{r}
nfr.dist_22 <- (data_transplants_aster%>%
  filter(year==2022))$n_tot_fr [(data_transplants_aster%>%
                             filter(year==2022))$fr_01==1]
length(nfr.dist_22) # 582
sum(nfr.dist_22 == 0) # 0 with fr_01==1 and nseed=0
nfr.parms_22 <- fitdistr(nfr.dist_22, "poisson")
nfr.parms_22 # lambda = 16.8419244

# Compare
hist(nfr.dist_22,col = rgb(1, 0, 0, 0.5),breaks=10)
hist(rpois(582, lambda = 16.8419244 ),col = rgb(0, 0, 1, 0.5),add=T,breaks=10)
```

Look at Poisson distributions for seeds:

```{r}
nseed.dist_22 <- (data_transplants_aster%>%
  filter(year==2022))$nseed[(data_transplants_aster%>%
                             filter(year==2022))$fr_01==1]
length(nseed.dist_22) # 582
sum(nseed.dist_22 == 0) # 0 with fr_01==1 and nseed=0
nseed.parms_22 <- fitdistr(nseed.dist_22, "poisson")
nseed.parms_22 # lambda = 85.7972509

# Compare
hist(nseed.dist_22,col = rgb(1, 0, 0, 0.5),breaks=10)
hist(rpois(582, lambda = 85.7972509),col = rgb(0, 0, 1, 0.5),add=T,breaks=10)
```

Look at negative binomial distributions for seeds:

```{r}
nseed.parms2_22 <- fitdistr(nseed.dist_22, "negative binomial")
# Warning message: In densfun(x, parm[1], parm[2], ...) : NaNs produced - OK?
nseed.parms2_22 # size = 3.9261436

# Compare
hist(nseed.dist_22,col = rgb(1, 0, 0, 0.5),breaks=10)
hist(rnbinom(582, size = 3.9261436, mu = 85.7972509),col = rgb(0, 0, 1, 0.5),
     add=T,breaks=10)

# NOT USED: parameters from distribution including plants with no seeds
# hist((data_transplants_aster%>%
#   filter(year==2022))$nseed,col = rgb(1, 0, 0, 0.5),breaks=10)
# hist(rnbinom(578, size = 0.31487453, mu = 58.88836105),col = rgb(0, 0, 1, 0.5),
#      add=T,breaks=10)
```

Specifying families:

```{r}
fam <- c(1,2,3,4) # specifies the one-parameter exponential families for the nodes
famlist <- list(fam.bernoulli(), # surv
                fam.bernoulli(), # fl_01
                fam.bernoulli(),# fr_01
                fam.truncated.negative.binomial(size=3.9261436,
                                                truncation=0)# nseed
                )
vars[fam == 1]
vars[fam == 2]
vars[fam == 3]
vars[fam == 4]
```

Specifying families (alternative, in case negative binomial really cannot be used with reaster):

```{r}
famlist1 <- list(fam.bernoulli(), # surv
                fam.bernoulli(), # fl_01
                fam.bernoulli(), # fr_01
                fam.truncated.poisson(truncation = 0) # nseed
                )
vars[fam == 1]
vars[fam == 2]
vars[fam == 3]
vars[fam == 4]
```

Creating the fitness variable of interest (and others):

```{r}
surv01 <- grep("surv", as.character(redata22$varb))
surv01 <- is.element(seq(along = redata22$varb), surv01)
redata22 <- data.frame(redata22, surv01 = as.integer(surv01))

fl01 <- grep("fl_01", as.character(redata22$varb))
fl01 <- is.element(seq(along = redata22$varb), fl01)
redata22 <- data.frame(redata22, fl01 = as.integer(fl01))

fr01 <- grep("fr_01", as.character(redata22$varb))
fr01 <- is.element(seq(along = redata22$varb), fr01)
redata22 <- data.frame(redata22, fr01 = as.integer(fr01))

seeds <- grep("nseed", as.character(redata22$varb))
seeds <- is.element(seq(along = redata22$varb), seeds)
redata22 <- data.frame(redata22, seeds = as.integer(seeds))

names(redata22)
```

Center the predictors (I think we do not need to use standardized predictors in Aster models, but centering the predictors helps the model to converge):

```{r}
redata22$temp_c<-redata22$temp-mean(redata22$temp)
redata22$temp_diff_c<-redata22$temp_diff-mean(redata22$temp_diff)
redata22$FFD_corr_c<-redata22$temp_diff-mean(redata22$FFD_corr)
```

Scaled predictors (in case we need to use them):

```{r}
redata22$temp_s<-scale(redata22$temp)
redata22$temp_diff_s<-scale(redata22$temp_diff)
redata22$FFD_corr_s<-scale(redata22$FFD_corr)
```

## 2023

Reshape data:

```{r}
redata23 <- reshape(data.frame(data_transplants_aster%>%
                                 filter(year==2023&!is.na(surv))%>%
                      dplyr::select(unique_id,plot,poll_type,temp,temp_diff,
                                    FFD_corr,surv,fl_01,fr_01,nseed)), 
                  varying = list(vars),
                  direction = "long", 
                  timevar = "varb", 
                  idvar = "unique_id",
                  times = as.factor(vars), v.names = "resp")
redata23<- data.frame(redata23, root = 1)
```

```{r}
names(data.frame(data_transplants_aster%>%filter(year==2023&!is.na(surv))%>%
                      dplyr::select(unique_id,plot,poll_type,temp,temp_diff,
                                    FFD_corr,surv,fl_01,fr_01,nseed)))
names(redata23)
nrow(data.frame(data_transplants_aster%>%filter(year==2023&!is.na(surv))%>%
                      dplyr::select(unique_id,plot,poll_type,temp,temp_diff,
                                    FFD_corr,surv,fl_01,fr_01,nseed)))
nrow(redata23)
nrow(data.frame(data_transplants_aster%>%filter(year==2023&!is.na(surv))%>%
                  dplyr::select(unique_id,plot,poll_type,temp,temp_diff,
                                FFD_corr,surv,fl_01,fr_01,nseed)))*length(vars)
sapply(redata23, class)
levels(redata23$varb)
length(unique(redata23$unique_id))
```

Look at Poisson distributions for fruits:

```{r}
nfr.dist_23 <- (data_transplants_aster%>%
  filter(year==2023))$n_tot_fr [(data_transplants_aster%>%
                             filter(year==2023))$fr_01==1]
length(nfr.dist_23) # 390
sum(nfr.dist_23 == 0) # 0 with fr_01==1 and nseed=0
nfr.parms_23 <- fitdistr(nfr.dist_23, "poisson")
nfr.parms_23 # lambda = 12.4282051

# Compare
hist(nfr.dist_23,col = rgb(1, 0, 0, 0.5),breaks=10)
hist(rpois(390, lambda = 12.4282051  ),col = rgb(0, 0, 1, 0.5),add=T,breaks=10)
```

Look at Poisson distributions for seeds:

```{r}
nseed.dist_23 <- (data_transplants_aster%>%
  filter(year==2023))$nseed[(data_transplants_aster%>%
                             filter(year==2023))$fr_01==1]
length(nseed.dist_23) # 390
sum(nseed.dist_23 == 0) # 0 with fr_01==1 and nseed=0
nseed.parms_23 <- fitdistr(nseed.dist_23, "poisson")
nseed.parms_23 # lambda = 62.5948718 

# Compare
hist(nseed.dist_23,col = rgb(1, 0, 0, 0.5),breaks=10)
hist(rpois(390, lambda = 62.5948718),col = rgb(0, 0, 1, 0.5),add=T,breaks=10)
```

Look at negative binomial distributions for flowers, fruits and seeds:

```{r}
nseed.parms2_23 <- fitdistr(nseed.dist_23, "negative binomial")
# Warning message: In densfun(x, parm[1], parm[2], ...) : NaNs produced - OK?
nseed.parms2_23 # size = 3.4636194

# Compare
hist(nseed.dist_23,col = rgb(1, 0, 0, 0.5),breaks=10)
hist(rnbinom(390, size = 3.4636194, mu = 62.5948718),col = rgb(0, 0, 1, 0.5),
     add=T,breaks=10)
```

Specifying families:

```{r}
fam <- c(1,2,3,4) # specifies the one-parameter exponential families for the nodes
famlist <- list(fam.bernoulli(), # surv
                fam.bernoulli(), # fl_01
                fam.bernoulli(),# fr_01
                fam.truncated.negative.binomial(size=3.4636194,
                                                truncation=0)# nseed
                )
vars[fam == 1]
vars[fam == 2]
vars[fam == 3]
vars[fam == 4]
```

Creating the fitness variable of interest (and others):

```{r}
surv01 <- grep("surv", as.character(redata23$varb))
surv01 <- is.element(seq(along = redata23$varb), surv01)
redata23 <- data.frame(redata23, surv01 = as.integer(surv01))

fl01 <- grep("fl_01", as.character(redata23$varb))
fl01 <- is.element(seq(along = redata23$varb), fl01)
redata23 <- data.frame(redata23, fl01 = as.integer(fl01))

fls <- grep("nfl", as.character(redata23$varb))
fls <- is.element(seq(along = redata23$varb), fls)
redata23 <- data.frame(redata23, fls = as.integer(fls))

fr01 <- grep("fr_01", as.character(redata23$varb))
fr01 <- is.element(seq(along = redata23$varb), fr01)
redata23 <- data.frame(redata23, fr01 = as.integer(fr01))

frs <- grep("nfr", as.character(redata23$varb))
frs <- is.element(seq(along = redata23$varb), frs)
redata23 <- data.frame(redata23, frs = as.integer(frs))

seeds <- grep("nseed", as.character(redata23$varb))
seeds <- is.element(seq(along = redata23$varb), seeds)
redata23 <- data.frame(redata23, seeds = as.integer(seeds))

names(redata23)
```

Center the predictors (I think we do not need to use standardized predictors in Aster models, but centering the predictors helps the model to converge):

```{r}
redata23$temp_c<-redata23$temp-mean(redata23$temp)
redata23$temp_diff_c<-redata23$temp_diff-mean(redata23$temp_diff)
redata23$FFD_corr_c<-redata23$temp_diff-mean(redata23$FFD_corr)
```

Scaled predictors (in case we need to use them):

```{r}
redata23$temp_s<-scale(redata23$temp)
redata23$temp_diff_s<-scale(redata23$temp_diff)
redata23$FFD_corr_s<-scale(redata23$FFD_corr)
```

## Appendix 1: Fit distributions to n seeds

```{r}
# Create two distributions
nseed_dist_22 <- nseed.dist_22
pois_dist_22 <- rpois(582, lambda = 85.7972509)
nbinom_dist_22<-rnbinom(582, size = 3.9261436, mu = 85.7972509)

nseed_dist_23 <- nseed.dist_23
pois_dist_23 <- rpois(390, lambda = 62.5948718)
nbinom_dist_23<-rnbinom(390, size = 3.4636194, mu = 62.5948718)

# Combine the distributions into one data frame
df_22_pois <- data.frame(
  value = c(nseed_dist_22, pois_dist_22),
  group = factor(c(rep("Actual distribution", length(nseed_dist_22)), 
                   rep("Poisson fit", length(pois_dist_22))))
)%>%mutate(year="2022")
df_22_nb <- data.frame(
  value = c(nseed_dist_22, nbinom_dist_22),
  group = factor(c(rep("Actual distribution", length(nseed_dist_22)), 
                   rep("Negative binomial fit", length(nbinom_dist_22))))
)%>%mutate(year="2022")

df_23_pois <- data.frame(
  value = c(nseed_dist_23, pois_dist_23),
  group = factor(c(rep("Actual distribution", length(nseed_dist_23)), 
                   rep("Poisson fit", length(pois_dist_23))))
)%>%mutate(year="2023")
df_23_nb <- data.frame(
  value = c(nseed_dist_23, nbinom_dist_23),
  group = factor(c(rep("Actual distribution", length(nseed_dist_23)), 
                   rep("Negative binomial fit", length(nbinom_dist_23))))
)%>%mutate(year="2023")

df_pois<-rbind(df_22_pois,df_23_pois)
df_nb<-rbind(df_22_nb,df_23_nb)

annotations_pois <- data.frame(
  year = unique(df_pois$year),  # Facet variable
  label = c("Lambda = 85.797", "Lambda = 62.595"),  # Text for each panel
  x = 210,  
  y = 240  
)

annotations_nb <- data.frame(
  year = unique(df_nb$year),  # Facet variable
  label = c("Size = 3.926, mu = 85.797", "Size = 3.464, mu = 62.595"),  # Adjust as needed
  x = 300, 
  y = 120   
)

# Plots with ggplot2
App1<-grid.arrange(
  ggplot(df_pois,aes(x=value,fill=group,color=group)) +
    geom_histogram(position="identity",bins=20,alpha=0.5) +
    my_theme_legend()+facet_wrap(~year)+theme(legend.position="top")+
    labs(x="Value",y="Count")+ggtitle("A)")+theme(legend.title=element_blank())+
    geom_text(data=annotations_pois,aes(x=x,y=y,label=label), 
              inherit.aes=FALSE,hjust=1,vjust=1,family="Times New Roman"),
  ggplot(df_nb,aes(x=value,fill=group,color=group)) +
    geom_histogram(position="identity",bins=20,alpha=0.5) +
    my_theme_legend()+facet_wrap(~year)+theme(legend.position="top")+
    labs(x="Value",y="Count")+ggtitle("B)")+theme(legend.title=element_blank())+
    geom_text(data=annotations_nb,aes(x=x,y=y,label=label), 
              inherit.aes=FALSE,hjust=1,vjust=1,family="Times New Roman"),
  ncol=1)
ggsave(file="output/figures/App1.tiff", plot=App1,width=6,height=8,dpi=300,
       compression="lzw")
```

# Aster both vars, plot as fixed (Table 2, Fig 2)

## 2022

### Fit model

Using negative binomial for seeds.

```{r}
aster22_nb_plotfi <- aster(resp ~ varb+seeds:(temp_s+temp_diff_s+
                                                I(temp_diff_s^2)+
                                                poll_type+plot), 
                           pred, fam, famlist=famlist, varb,
                           idvar=as.numeric(unique_id), 
                           root, data = redata22)
```

```{r}
summary(aster22_nb_plotfi)
```

### Table 2: LRTs

Testing the significance of seeds:temp_s

```{r}
aster22_nb_plotfi_1 <- aster(resp ~ varb+seeds:(temp_diff_s+
                                                  I(temp_diff_s^2)+
                                                  poll_type+plot), 
                             pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                             root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_1,aster22_nb_plotfi)
```

Testing the significance of seeds:temp_diff_s

```{r}
aster22_nb_plotfi_2 <- aster(resp ~ varb+seeds:(temp_s+
                                                  I(temp_diff_s^2)+
                                                  poll_type+plot), 
                             pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                             root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_2,aster22_nb_plotfi)
```

Testing the significance of seeds:I(temp_diff_s^2)

```{r}
aster22_nb_plotfi_3 <- aster(resp ~ varb+seeds:(temp_s+temp_diff_s+
                                                  poll_type+plot), 
                             pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                             root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_3,aster22_nb_plotfi)
```

Testing the significance of seeds:poll_type

```{r}
aster22_nb_plotfi_4 <- aster(resp ~ varb+seeds:(temp_s+temp_diff_s+
                                                  I(temp_diff_s^2)+
                                                  plot), 
                             pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                             root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_4,aster22_nb_plotfi)
```

Testing the significance of seeds:plot

```{r}
aster22_nb_plotfi_5 <- aster(resp ~ varb+seeds:(temp_s+temp_diff_s+
                                                  I(temp_diff_s^2)+
                                                  poll_type), 
                             pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                             root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_5,aster22_nb_plotfi)
```

### Predictions

##### seeds:temp_s (NS)

```{r}
# Predictions calculated for values ranging from the 5th percentile to the 95th 
# percentile, to avoid making predictions based on very few observations
# NOT DONE HERE! Using min and max.

# make prediction df
aster22_plotfi_predict1 <- data.frame(
  expand_grid(temp_s = seq(from = as.numeric(range(redata22$temp_s)[1]),
                           to = as.numeric(range(redata22$temp_s)[2]),
                           length.out = 100)),
  #OK from here to below? Not sure...
  temp_diff_s = 0, # scaled so mean=0
  surv = 1,
  fl_01 = 1,
  fr_01 = 1,
  nseed = 0,
  root = 1
)
aster22_plotfi_predict1<-rbind(aster22_plotfi_predict1%>%
                                 mutate(poll_type=as.factor("Crosspollinated")),
                               aster22_plotfi_predict1%>%
                                 mutate(poll_type=as.factor("Selfpollinated")))
aster22_plotfi_predict1<-rbind(aster22_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[1])),
                               aster22_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[2])),
                               aster22_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[3])),
                               aster22_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[4])),
                               aster22_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[5])),
                               aster22_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[6])),
                               aster22_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[7])),
                               aster22_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[8])))

# reshape data to long format
aster22_plotfi_predict1_long <- reshape(as.data.frame(aster22_plotfi_predict1), 
                                        varying = list(vars), direction = "long", 
                                        timevar = "varb", times = as.factor(vars), 
                                        v.names = "resp")

# add artifice
aster22_plotfi_predict1_long$seeds <- as.integer(ifelse(
  aster22_plotfi_predict1_long$varb == "nseed", 1, 0))

aster22_plotfi_predict1.df <- aster22_plotfi_predict1_long

# fixed model predictions
aster22_plotfi.p1 <- predict(aster22_nb_plotfi,
                             newdata = aster22_plotfi_predict1_long,
                             varvar = varb,idvar = id, root = root,
                             #info.tol = 1e-11, 
                             se.fit = T)

aster22_plotfi_predict1.df$fit <- aster22_plotfi.p1$fit
aster22_plotfi_predict1.df$se <- aster22_plotfi.p1$se.fit

aster22_plotfi_predict1.df.seeds <- aster22_plotfi_predict1.df %>%
  filter(varb == "nseed")

aster22_plotfi_predict1.df.seeds<-aster22_plotfi_predict1.df.seeds%>%
  mutate(ymin=fit-se,ymax=fit+se)

original_mean_temp_22<-as.numeric(data_transplants_aster%>%
                                    filter(year==2022)%>%
                                    summarise(mean_temp=mean(temp)))
original_sd_temp_22<-as.numeric(data_transplants_aster%>%
                                  filter(year==2022)%>%
                                  summarise(sd_temp=sd(temp)))

aster22_plotfi_predict1.df.seeds<-aster22_plotfi_predict1.df.seeds%>%
  mutate(temp=temp_s*original_sd_temp_22+original_mean_temp_22)

aster22_plotfi_predict1.df.seeds<-aster22_plotfi_predict1.df.seeds%>%
  group_by(temp)%>%
  summarise(fit=mean(fit),se=mean(se),ymin=mean(ymin),ymax=mean(ymax))
```

#### seeds:temp_diff_s (*)

```{r}
# make prediction df
aster22_plotfi_predict2 <- data.frame(
  expand_grid(temp_diff_s = seq(from = as.numeric(range(redata22$temp_diff_s)[1]),
                                to = as.numeric(range(redata22$temp_diff_s)[2]),
                                length.out = 100)),
  #OK from here to below? Not sure...
  temp_s = 0, # scaled so mean=0
  surv = 1,
  fl_01 = 1,
  fr_01 = 1,
  nseed = 0,
  root = 1
)

aster22_plotfi_predict2<-rbind(aster22_plotfi_predict2%>%
                                 mutate(poll_type=as.factor("Crosspollinated")),
                               aster22_plotfi_predict2%>%
                                 mutate(poll_type=as.factor("Selfpollinated")))
aster22_plotfi_predict2<-rbind(aster22_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[1])),
                               aster22_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[2])),
                               aster22_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[3])),
                               aster22_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[4])),
                               aster22_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[5])),
                               aster22_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[6])),
                               aster22_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[7])),
                               aster22_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[8])))

# reshape data to long format
aster22_plotfi_predict2_long <- reshape(as.data.frame(aster22_plotfi_predict2), 
                                        varying = list(vars), direction = "long", 
                                        timevar = "varb", times = as.factor(vars), 
                                        v.names = "resp")

# add artifice
aster22_plotfi_predict2_long$seeds <- as.integer(ifelse(
  aster22_plotfi_predict2_long$varb == "nseed", 1, 0))

aster22_plotfi_predict2.df <- aster22_plotfi_predict2_long

# fixed model predictions
aster22_plotfi.p2 <- predict(aster22_nb_plotfi, # Use aster model and not reaster!
                             newdata = aster22_plotfi_predict2_long,
                             varvar = varb,idvar = id, root = root,
                             #info.tol = 1e-11, 
                             se.fit = T)

aster22_plotfi_predict2.df$fit <- aster22_plotfi.p2$fit
aster22_plotfi_predict2.df$se <- aster22_plotfi.p2$se.fit

aster22_plotfi_predict2.df.seeds <- aster22_plotfi_predict2.df %>%
  filter(varb == "nseed")

aster22_plotfi_predict2.df.seeds<-aster22_plotfi_predict2.df.seeds%>%
  mutate(ymin=fit-se,ymax=fit+se)

original_mean_temp_diff_22<-as.numeric(data_transplants_aster%>%
                                         filter(year==2022)%>%
                                         summarise(mean_temp_diff=mean(temp_diff)))
original_sd_temp_diff_22<-as.numeric(data_transplants_aster%>%
                                       filter(year==2022)%>%
                                       summarise(sd_temp_diff=sd(temp_diff)))

aster22_plotfi_predict2.df.seeds<-aster22_plotfi_predict2.df.seeds%>%
  mutate(temp_diff=temp_diff_s*original_sd_temp_diff_22+
           original_mean_temp_diff_22)

aster22_plotfi_predict2.df.seeds<-aster22_plotfi_predict2.df.seeds%>%
  group_by(temp_diff)%>%
  summarise(fit=mean(fit),se=mean(se),ymin=mean(ymin),ymax=mean(ymax))
```

## 2023

### Fit model

Using negative binomial for seeds.

```{r}
aster23_nb_plotfi <- aster(resp ~ varb+seeds:(temp_s+temp_diff_s+
                                                I(temp_diff_s^2)+
                                                poll_type+plot), 
                           pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                           root, data = redata23)
```

```{r}
summary(aster23_nb_plotfi,info.tol=1e-15)
```

### Table 2: LRTs

Testing the significance of seeds:temp_s

```{r}
aster23_nb_plotfi_1 <- aster(resp ~ varb+seeds:(temp_diff_s+
                                                  I(temp_diff_s^2)+
                                                  poll_type+plot), 
                             pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                             root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_1,aster23_nb_plotfi)
```

Testing the significance of seeds:temp_diff_s

```{r}
aster23_nb_plotfi_2 <- aster(resp ~ varb+seeds:(temp_s+
                                                  I(temp_diff_s^2)+
                                                  poll_type+plot), 
                             pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                             root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_2,aster23_nb_plotfi)
```

Testing the significance of seeds:I(temp_diff_s^2)

```{r}
aster23_nb_plotfi_3 <- aster(resp ~ varb+seeds:(temp_s+temp_diff_s+
                                                  poll_type+plot), 
                             pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                             root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_3,aster23_nb_plotfi)
```

Testing the significance of seeds:poll_type

```{r}
aster23_nb_plotfi_4 <- aster(resp ~ varb+seeds:(temp_s+temp_diff_s+
                                                  I(temp_diff_s^2)+
                                                  plot), 
                             pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                             root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_4,aster23_nb_plotfi)
```

Testing the significance of seeds:plot

```{r}
aster23_nb_plotfi_5 <- aster(resp ~ varb+seeds:(temp_s+temp_diff_s+
                                                  I(temp_diff_s^2)+
                                                  poll_type), 
                             pred, fam, famlist=famlist, varb, idvar=as.numeric(unique_id), 
                             root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_5,aster23_nb_plotfi)
```

### Predictions

##### seeds:temp_s (*)

```{r}
# Predictions calculated for values ranging from the 5th percentile to the 95th 
# percentile, to avoid making predictions based on very few observations
# NOT DONE HERE! Using min and max.

# make prediction df
aster23_plotfi_predict1 <- data.frame(
  expand_grid(temp_s = seq(from = as.numeric(range(redata23$temp_s)[1]),
                           to = as.numeric(range(redata23$temp_s)[2]),
                           length.out = 100)),
  #OK from here to below? Not sure...
  temp_diff_s = 0, # scaled so mean=0
  surv = 1,
  fl_01 = 1,
  fr_01 = 1,
  nseed = 0,
  root = 1
)
aster23_plotfi_predict1<-rbind(aster23_plotfi_predict1%>%
                                 mutate(poll_type=as.factor("Crosspollinated")),
                               aster23_plotfi_predict1%>%
                                 mutate(poll_type=as.factor("Selfpollinated")))
aster23_plotfi_predict1<-rbind(aster23_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[1])),
                               aster23_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[2])),
                               aster23_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[3])),
                               aster23_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[4])),
                               aster23_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[5])),
                               aster23_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[6])),
                               aster23_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[7])),
                               aster23_plotfi_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[8])))

# reshape data to long format
aster23_plotfi_predict1_long <- reshape(as.data.frame(aster23_plotfi_predict1), 
                                        varying = list(vars), direction = "long", 
                                        timevar = "varb", times = as.factor(vars), 
                                        v.names = "resp")

# add artifice
aster23_plotfi_predict1_long$seeds <- as.integer(ifelse(
  aster23_plotfi_predict1_long$varb == "nseed", 1, 0))

aster23_plotfi_predict1.df <- aster23_plotfi_predict1_long

# fixed model predictions
aster23_plotfi.p1 <- predict(aster23_nb_plotfi,
                             newdata = aster23_plotfi_predict1_long,
                             varvar = varb,idvar = id, root = root,
                             info.tol = 1e-15, 
                             se.fit = T)

aster23_plotfi_predict1.df$fit <- aster23_plotfi.p1$fit
aster23_plotfi_predict1.df$se <- aster23_plotfi.p1$se.fit

aster23_plotfi_predict1.df.seeds <- aster23_plotfi_predict1.df %>%
  filter(varb == "nseed")

aaster23_plotfi_predict1.df.seeds<-aster23_plotfi_predict1.df.seeds%>%
  mutate(ymin=fit-se,ymax=fit+se)

original_mean_temp_23<-as.numeric(data_transplants_aster%>%
                                    filter(year==2023)%>%
                                    summarise(mean_temp=mean(temp)))
original_sd_temp_23<-as.numeric(data_transplants_aster%>%
                                  filter(year==2023)%>%
                                  summarise(sd_temp=sd(temp)))

aaster23_plotfi_predict1.df.seeds<-aaster23_plotfi_predict1.df.seeds%>%
  mutate(temp=temp_s*original_sd_temp_23+original_mean_temp_23)

aaster23_plotfi_predict1.df.seeds<-aaster23_plotfi_predict1.df.seeds%>%
  group_by(temp)%>%
  summarise(fit=mean(fit),se=mean(se),ymin=mean(ymin),ymax=mean(ymax))
```

#### seeds:temp_diff_s (NS - almost)

```{r}
# make prediction df
aster23_plotfi_predict2 <- data.frame(
  expand_grid(temp_diff_s = seq(from = as.numeric(range(redata23$temp_diff_s)[1]),
                                to = as.numeric(range(redata23$temp_diff_s)[2]),
                                length.out = 100)),
  #OK from here to below? Not sure...
  temp_s = 0, # scaled so mean=0
  surv = 1,
  fl_01 = 1,
  fr_01 = 1,
  nseed = 0,
  root = 1
)

aster23_plotfi_predict2<-rbind(aster23_plotfi_predict2%>%
                                 mutate(poll_type=as.factor("Crosspollinated")),
                               aster23_plotfi_predict2%>%
                                 mutate(poll_type=as.factor("Selfpollinated")))
aster23_plotfi_predict2<-rbind(aster23_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[1])),
                               aster23_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[2])),
                               aster23_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[3])),
                               aster23_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[4])),
                               aster23_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[5])),
                               aster23_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[6])),
                               aster23_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[7])),
                               aster23_plotfi_predict2%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[8])))

# reshape data to long format
aster23_plotfi_predict2_long <- reshape(as.data.frame(aster23_plotfi_predict2), 
                                        varying = list(vars), direction = "long", 
                                        timevar = "varb", times = as.factor(vars), 
                                        v.names = "resp")

# add artifice
aster23_plotfi_predict2_long$seeds <- as.integer(ifelse(
  aster23_plotfi_predict2_long$varb == "nseed", 1, 0))

aster23_plotfi_predict2.df <- aster23_plotfi_predict2_long

# fixed model predictions
aster23_plotfi.p2 <- predict(aster23_nb_plotfi, # Use aster model and not reaster!
                             newdata = aster23_plotfi_predict2_long,
                             varvar = varb,idvar = id, root = root,
                             info.tol = 1e-15, 
                             se.fit = T)

aster23_plotfi_predict2.df$fit <- aster23_plotfi.p2$fit
aster23_plotfi_predict2.df$se <- aster23_plotfi.p2$se.fit

aster23_plotfi_predict2.df.seeds <- aster23_plotfi_predict2.df %>%
  filter(varb == "nseed")

aster23_plotfi_predict2.df.seeds<-aster23_plotfi_predict2.df.seeds%>%
  mutate(ymin=fit-se,ymax=fit+se)

original_mean_temp_diff_23<-as.numeric(data_transplants_aster%>%
                                         filter(year==2023)%>%
                                         summarise(mean_temp_diff=mean(temp_diff)))
original_sd_temp_diff_23<-as.numeric(data_transplants_aster%>%
                                       filter(year==2023)%>%
                                       summarise(sd_temp_diff=sd(temp_diff)))

aster23_plotfi_predict2.df.seeds<-aster23_plotfi_predict2.df.seeds%>%
  mutate(temp_diff=temp_diff_s*original_sd_temp_diff_23+
           original_mean_temp_diff_23)

aster23_plotfi_predict2.df.seeds<-aster23_plotfi_predict2.df.seeds%>%
  group_by(temp_diff)%>%
  summarise(fit=mean(fit),se=mean(se),ymin=mean(ymin),ymax=mean(ymax))
```

## Figure 2

```{r}
fig2<-grid.arrange(
  ggplot()+
    geom_point(data=data_transplants%>%filter(fruiting==1),
               aes(x=temp,y=n_tot_seed,color=year),size=1,alpha=0.2)+
    # 2022 temp
    geom_ribbon(data=aster22_plotfi_predict1.df.seeds,
                aes(x=temp,y=fit,ymin=ymin,ymax=ymax),
                fill="#F8766D",alpha=0.25)+
    geom_line(data=aster22_plotfi_predict1.df.seeds,
              aes(x=temp,y=fit),size=0.5,color="#F8766D",linetype="dashed")+
    # 2023 temp
    geom_ribbon(data=aaster23_plotfi_predict1.df.seeds,
                aes(x=temp,y=fit,ymin=ymin,ymax=ymax),
                fill="#00BFC4",alpha=0.25)+
    geom_line(data=aaster23_plotfi_predict1.df.seeds,
              aes(x=temp,y=fit),size=0.5,color="#00BFC4")+
    xlab("Temperature at planting site (ºC)")+ylab(NULL)+
    my_theme(),
  ggplot()+
    geom_point(data=data_transplants%>%filter(fruiting==1),
               aes(x=temp_diff,y=n_tot_seed,color=year),size=1,alpha=0.2)+
    # 2022 temp_diff
    geom_ribbon(data=aster22_plotfi_predict2.df.seeds,
                aes(x=temp_diff,y=fit,ymin=ymin,ymax=ymax),
                fill="#F8766D",alpha=0.25)+
    geom_line(data=aster22_plotfi_predict2.df.seeds,
              aes(x=temp_diff,y=fit),size=0.5,color="#F8766D")+
    # 2023 temp_diff
    geom_ribbon(data=aster23_plotfi_predict2.df.seeds,
                aes(x=temp_diff,y=fit,ymin=ymin,ymax=ymax),
                fill="#00BFC4",alpha=0.25)+
    geom_line(data=aster23_plotfi_predict2.df.seeds,
              aes(x=temp_diff,y=fit),size=0.5,color="#00BFC4",linetype="dashed")+
    xlab("Temperature difference (ºC)")+ylab(NULL)+
    my_theme(),
  ncol=2,left=textGrob("     Predicted fitness",rot=90,
                       gp=gpar(fontsize=16,fontfamily="serif"),just="center"))
ggsave(file="output/figures/fig2.tiff", plot=fig2,width=7,height=3.5,dpi=300,
       compression="lzw")
```

```{r}
fig2_nopoints<-grid.arrange(
  ggplot()+
    #geom_point(data=data_transplants%>%filter(fruiting==1),
    #           aes(x=temp,y=n_tot_seed,color=year),size=1.5,alpha=0.2)+
    # 2022 temp
    geom_ribbon(data=aster22_plotfi_predict1.df.seeds,
                aes(x=temp,y=fit,ymin=ymin,ymax=ymax),
                fill="#F8766D",alpha=0.25)+
    geom_line(data=aster22_plotfi_predict1.df.seeds,
              aes(x=temp,y=fit),size=0.5,color="#F8766D",linetype="dashed")+
    # 2023 temp
    geom_ribbon(data=aaster23_plotfi_predict1.df.seeds,
                aes(x=temp,y=fit,ymin=ymin,ymax=ymax),
                fill="#00BFC4",alpha=0.25)+
    geom_line(data=aaster23_plotfi_predict1.df.seeds,
              aes(x=temp,y=fit),size=0.5,color="#00BFC4")+
    xlab("Temperature at planting site (ºC)")+ylab(NULL)+
    my_theme()+scale_y_continuous(limits=c(0,85)),
  ggplot()+
    #geom_point(data=data_transplants%>%filter(fruiting==1),
    #           aes(x=temp_diff,y=n_tot_seed,color=year),size=1.5,alpha=0.2)+
    # 2022 temp_diff
    geom_ribbon(data=aster22_plotfi_predict2.df.seeds,
                aes(x=temp_diff,y=fit,ymin=ymin,ymax=ymax),
                fill="#F8766D",alpha=0.25)+
    geom_line(data=aster22_plotfi_predict2.df.seeds,
              aes(x=temp_diff,y=fit),size=0.5,color="#F8766D")+
    # 2023 temp_diff
    geom_ribbon(data=aster23_plotfi_predict2.df.seeds,
                aes(x=temp_diff,y=fit,ymin=ymin,ymax=ymax),
                fill="#00BFC4",alpha=0.25)+
    geom_line(data=aster23_plotfi_predict2.df.seeds,
              aes(x=temp_diff,y=fit),size=0.5,color="#00BFC4",linetype="dashed")+
    xlab("Temperature difference (ºC)")+ylab(NULL)+
    my_theme()+scale_y_continuous(limits=c(0,85)),
  ncol=2,left=textGrob("    Predicted fitness",rot=90,
                       gp=gpar(fontsize=16,fontfamily="serif"),just="center"))
ggsave(file="output/figures/fig2_nopoints.tiff",
       plot=fig2_nopoints,width=7,height=3.5,dpi=300,compression="lzw")
```

# Aster single vars, plot as fixed (App 6-7)

## temp

### 2022

#### Fit model

```{r}
aster22_nb_plotfi_temp <- aster(resp ~ varb+seeds:(temp_s+poll_type+plot),
                                   pred, fam, famlist=famlist, varb,
                                   idvar=as.numeric(unique_id), 
                                   root, data = redata22)
```

```{r}
summary(aster22_nb_plotfi_temp)
```

#### Appendix 6: LRTs

Submodels:

Testing the significance of seeds:temp_s

```{r}
aster22_nb_plotfi_temp_1 <- aster(resp ~ varb+seeds:(poll_type+plot),
                                   pred, fam, famlist=famlist, varb,
                                   idvar=as.numeric(unique_id), 
                                   root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_temp_1,aster22_nb_plotfi_temp)
```

Testing the significance of seeds:poll_type

```{r}
aster22_nb_plotfi_temp_2 <- aster(resp ~ varb+seeds:(temp_s+plot),
                                   pred, fam, famlist=famlist, varb,
                                   idvar=as.numeric(unique_id), 
                                   root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_temp_2,aster22_nb_plotfi_temp)
```

Testing the significance of seeds:plot

```{r}
aster22_nb_plotfi_temp_3 <- aster(resp ~ varb+seeds:(temp_s+poll_type),
                                   pred, fam, famlist=famlist, varb,
                                   idvar=as.numeric(unique_id), 
                                   root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_temp_3,aster22_nb_plotfi_temp)
```

#### Predictions

```{r}
# Predictions calculated for values ranging from the 5th percentile to the 95th 
# percentile, to avoid making predictions based on very few observations
# NOT DONE HERE! Using min and max.

# make prediction df
aster22_nb_plotfi_temp_predict1 <- data.frame(
  expand_grid(temp_s = seq(from = as.numeric(range(redata22$temp_s)[1]),
                           to = as.numeric(range(redata22$temp_s)[2]),
                           length.out = 100)),
  #OK from here to below? Not sure...
  temp_diff_s = 0, # scaled so mean=0
  surv = 1,
  fl_01 = 1,
  fr_01 = 1,
  nseed = 0,
  root = 1
)
aster22_nb_plotfi_temp_predict1<-rbind(aster22_nb_plotfi_temp_predict1%>%
                                         mutate(poll_type=as.factor("Crosspollinated")),
                                       aster22_nb_plotfi_temp_predict1%>%
                                         mutate(poll_type=as.factor("Selfpollinated")))
aster22_nb_plotfi_temp_predict1<-rbind(aster22_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[1])),
                               aster22_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[2])),
                               aster22_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[3])),
                               aster22_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[4])),
                               aster22_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[5])),
                               aster22_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[6])),
                               aster22_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[7])),
                               aster22_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[8])))

# reshape data to long format
aster22_nb_plotfi_temp_predict1_long <- reshape(as.data.frame(aster22_nb_plotfi_temp_predict1), 
                                        varying = list(vars), direction = "long", 
                                        timevar = "varb", times = as.factor(vars), 
                                        v.names = "resp")

# add artifice
aster22_nb_plotfi_temp_predict1_long$seeds <- as.integer(ifelse(
  aster22_nb_plotfi_temp_predict1_long$varb == "nseed", 1, 0))

aster22_nb_plotfi_temp_predict1.df <- aster22_nb_plotfi_temp_predict1_long

# fixed model predictions
aster22_nb_plotfi_temp.p1 <- predict(aster22_nb_plotfi_temp,
                             newdata = aster22_nb_plotfi_temp_predict1_long,
                             varvar = varb,idvar = id, root = root,
                             #info.tol = 1e-11, 
                             se.fit = T)

aster22_nb_plotfi_temp_predict1.df$fit <- aster22_nb_plotfi_temp.p1$fit
aster22_nb_plotfi_temp_predict1.df$se <- aster22_nb_plotfi_temp.p1$se.fit

aster22_nb_plotfi_temp_predict1.df.seeds <- aster22_nb_plotfi_temp_predict1.df %>%
  filter(varb == "nseed")

aster22_nb_plotfi_temp_predict1.df.seeds<-aster22_nb_plotfi_temp_predict1.df.seeds%>%
  mutate(ymin=fit-se,ymax=fit+se)

aster22_nb_plotfi_temp_predict1.df.seeds<-aster22_nb_plotfi_temp_predict1.df.seeds%>%
  mutate(temp=temp_s*original_sd_temp_22+original_mean_temp_22)

aster22_nb_plotfi_temp_predict1.df.seeds<-aster22_nb_plotfi_temp_predict1.df.seeds%>%
  group_by(temp)%>%
  summarise(fit=mean(fit),se=mean(se),ymin=mean(ymin),ymax=mean(ymax))
```

### 2023

#### Fit model

```{r}
aster23_nb_plotfi_temp <- aster(resp ~ varb+seeds:(temp_s+poll_type+plot),
                                   pred, fam, famlist=famlist, varb,
                                   idvar=as.numeric(unique_id), 
                                   root, data = redata23)
```

```{r}
summary(aster23_nb_plotfi_temp,info.tol=1e-15)
```

#### Appendix 6: LRTs

Submodels:

Testing the significance of seeds:temp_s

```{r}
aster23_nb_plotfi_temp_1 <- aster(resp ~ varb+seeds:(poll_type+plot),
                                   pred, fam, famlist=famlist, varb,
                                   idvar=as.numeric(unique_id), 
                                   root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_temp_1,aster23_nb_plotfi_temp)
```

Testing the significance of seeds:poll_type

```{r}
aster23_nb_plotfi_temp_2 <- aster(resp ~ varb+seeds:(temp_s+plot),
                                   pred, fam, famlist=famlist, varb,
                                   idvar=as.numeric(unique_id), 
                                   root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_temp_2,aster23_nb_plotfi_temp)
```

Testing the significance of seeds:plot

```{r}
aster23_nb_plotfi_temp_3 <- aster(resp ~ varb+seeds:(temp_s+poll_type),
                                   pred, fam, famlist=famlist, varb,
                                   idvar=as.numeric(unique_id), 
                                   root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_temp_3,aster23_nb_plotfi_temp)
```

#### Predictions

```{r}
# Predictions calculated for values ranging from the 5th percentile to the 95th 
# percentile, to avoid making predictions based on very few observations
# NOT DONE HERE! Using min and max.

# make prediction df
aster23_nb_plotfi_temp_predict1 <- data.frame(
  expand_grid(temp_s = seq(from = as.numeric(range(redata23$temp_s)[1]),
                           to = as.numeric(range(redata23$temp_s)[2]),
                           length.out = 100)),
  #OK from here to below? Not sure...
  temp_diff_s = 0, # scaled so mean=0
  surv = 1,
  fl_01 = 1,
  fr_01 = 1,
  nseed = 0,
  root = 1
)
aster23_nb_plotfi_temp_predict1<-rbind(aster23_nb_plotfi_temp_predict1%>%
                                         mutate(poll_type=as.factor("Crosspollinated")),
                                       aster23_nb_plotfi_temp_predict1%>%
                                         mutate(poll_type=as.factor("Selfpollinated")))
aster23_nb_plotfi_temp_predict1<-rbind(aster23_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[1])),
                               aster23_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[2])),
                               aster23_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[3])),
                               aster23_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[4])),
                               aster23_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[5])),
                               aster23_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[6])),
                               aster23_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[7])),
                               aster23_nb_plotfi_temp_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[8])))

# reshape data to long format
aster23_nb_plotfi_temp_predict1_long <- reshape(as.data.frame(aster23_nb_plotfi_temp_predict1), 
                                        varying = list(vars), direction = "long", 
                                        timevar = "varb", times = as.factor(vars), 
                                        v.names = "resp")

# add artifice
aster23_nb_plotfi_temp_predict1_long$seeds <- as.integer(ifelse(
  aster23_nb_plotfi_temp_predict1_long$varb == "nseed", 1, 0))

aster23_nb_plotfi_temp_predict1.df <- aster23_nb_plotfi_temp_predict1_long

# fixed model predictions
aster23_nb_plotfi_temp.p1 <- predict(aster23_nb_plotfi_temp,
                             newdata = aster23_nb_plotfi_temp_predict1_long,
                             varvar = varb,idvar = id, root = root,
                             info.tol = 1e-15, 
                             se.fit = T)

aster23_nb_plotfi_temp_predict1.df$fit <- aster23_nb_plotfi_temp.p1$fit
aster23_nb_plotfi_temp_predict1.df$se <- aster23_nb_plotfi_temp.p1$se.fit

aster23_nb_plotfi_temp_predict1.df.seeds <- aster23_nb_plotfi_temp_predict1.df %>%
  filter(varb == "nseed")

aster23_nb_plotfi_temp_predict1.df.seeds<-aster23_nb_plotfi_temp_predict1.df.seeds%>%
  mutate(ymin=fit-se,ymax=fit+se)

aster23_nb_plotfi_temp_predict1.df.seeds<-aster23_nb_plotfi_temp_predict1.df.seeds%>%
  mutate(temp=temp_s*original_sd_temp_23+original_mean_temp_23)

aster23_nb_plotfi_temp_predict1.df.seeds<-aster23_nb_plotfi_temp_predict1.df.seeds%>%
  group_by(temp)%>%
  summarise(fit=mean(fit),se=mean(se),ymin=mean(ymin),ymax=mean(ymax))
```

## temp_diff

### 2022

#### Fit model

```{r}
aster22_nb_plotfi_temp_diff <- aster(resp ~ varb+seeds:(temp_diff_s+
                                                          I(temp_diff_s^2)+
                                                          poll_type+plot), 
                                  pred, fam, famlist=famlist, varb,
                                  idvar=as.numeric(unique_id), 
                                  root, data = redata22)
```

```{r}
summary(aster22_nb_plotfi_temp_diff)
```

#### Appendix 6: LRTs

Submodels:

Testing the significance of seeds:temp_diff_s

```{r}
aster22_nb_plotfi_temp_diff_1 <- aster(resp ~ varb+seeds:(I(temp_diff_s^2)+
                                                          poll_type+plot),
                                   pred, fam, famlist=famlist, varb,
                                   idvar=as.numeric(unique_id), 
                                   root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_temp_diff_1,aster22_nb_plotfi_temp_diff)
```

Testing the significance of seeds:I(temp_diff^2)

```{r}
aster22_nb_plotfi_temp_diff_2 <- aster(resp ~ varb+seeds:(temp_diff_s+
                                                          poll_type+plot), 
                                  pred, fam, famlist=famlist, varb,
                                  idvar=as.numeric(unique_id), 
                                  root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_temp_diff_2,aster22_nb_plotfi_temp_diff)
```

Testing the significance of seeds:poll_type

```{r}
aster22_nb_plotfi_temp_diff_3 <- aster(resp ~ varb+seeds:(temp_diff_s+
                                                          I(temp_diff_s^2)+
                                                          plot), 
                                  pred, fam, famlist=famlist, varb,
                                  idvar=as.numeric(unique_id), 
                                  root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_temp_diff_3,aster22_nb_plotfi_temp_diff)
```

Testing the significance of seeds:plot

```{r}
aster22_nb_plotfi_temp_diff_4 <- aster(resp ~ varb+seeds:(temp_diff_s+
                                                          I(temp_diff_s^2)+
                                                          poll_type), 
                                  pred, fam, famlist=famlist, varb,
                                  idvar=as.numeric(unique_id), 
                                  root, data = redata22)
```

```{r}
anova(aster22_nb_plotfi_temp_diff_4,aster22_nb_plotfi_temp_diff)
```

#### Predictions

```{r}
# Predictions calculated for values ranging from the 5th percentile to the 95th 
# percentile, to avoid making predictions based on very few observations
# NOT DONE HERE! Using min and max.

# make prediction df
aster22_nb_plotfi_temp_diff_predict1 <- data.frame(
  expand_grid(temp_diff_s = seq(from = as.numeric(range(redata22$temp_diff_s)[1]),
                           to = as.numeric(range(redata22$temp_diff_s)[2]),
                           length.out = 100)),
  #OK from here to below? Not sure...
  temp_s = 0, # scaled so mean=0
  surv = 1,
  fl_01 = 1,
  fr_01 = 1,
  nseed = 0,
  root = 1
)
aster22_nb_plotfi_temp_diff_predict1<-rbind(aster22_nb_plotfi_temp_diff_predict1%>%
                                         mutate(poll_type=as.factor("Crosspollinated")),
                                       aster22_nb_plotfi_temp_diff_predict1%>%
                                         mutate(poll_type=as.factor("Selfpollinated")))
aster22_nb_plotfi_temp_diff_predict1<-rbind(aster22_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[1])),
                               aster22_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[2])),
                               aster22_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[3])),
                               aster22_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[4])),
                               aster22_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[5])),
                               aster22_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[6])),
                               aster22_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[7])),
                               aster22_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata22$plot)[8])))

# reshape data to long format
aster22_nb_plotfi_temp_diff_predict1_long <- reshape(as.data.frame(aster22_nb_plotfi_temp_diff_predict1), 
                                        varying = list(vars), direction = "long", 
                                        timevar = "varb", times = as.factor(vars), 
                                        v.names = "resp")

# add artifice
aster22_nb_plotfi_temp_diff_predict1_long$seeds <- as.integer(ifelse(
  aster22_nb_plotfi_temp_diff_predict1_long$varb == "nseed", 1, 0))

aster22_nb_plotfi_temp_diff_predict1.df <- aster22_nb_plotfi_temp_diff_predict1_long

# fixed model predictions
aster22_nb_plotfi_temp_diff.p1 <- predict(aster22_nb_plotfi_temp_diff,
                             newdata = aster22_nb_plotfi_temp_diff_predict1_long,
                             varvar = varb,idvar = id, root = root,
                             #info.tol = 1e-11, 
                             se.fit = T)

aster22_nb_plotfi_temp_diff_predict1.df$fit <- aster22_nb_plotfi_temp_diff.p1$fit
aster22_nb_plotfi_temp_diff_predict1.df$se <- aster22_nb_plotfi_temp_diff.p1$se.fit

aster22_nb_plotfi_temp_diff_predict1.df.seeds <- aster22_nb_plotfi_temp_diff_predict1.df %>%
  filter(varb == "nseed")

aster22_nb_plotfi_temp_diff_predict1.df.seeds<-aster22_nb_plotfi_temp_diff_predict1.df.seeds%>%
  mutate(ymin=fit-se,ymax=fit+se)

aster22_nb_plotfi_temp_diff_predict1.df.seeds<-aster22_nb_plotfi_temp_diff_predict1.df.seeds%>%
  mutate(temp_diff=temp_diff_s*original_sd_temp_diff_22+original_mean_temp_diff_22)

aster22_nb_plotfi_temp_diff_predict1.df.seeds<-aster22_nb_plotfi_temp_diff_predict1.df.seeds%>%
  group_by(temp_diff)%>%
  summarise(fit=mean(fit),se=mean(se),ymin=mean(ymin),ymax=mean(ymax))
```

### 2023

#### Fit model

```{r}
aster23_nb_plotfi_temp_diff <- aster(resp ~ varb+seeds:(temp_diff_s+
                                                          I(temp_diff_s^2)+
                                                          poll_type+plot), 
                                  pred, fam, famlist=famlist, varb,
                                  idvar=as.numeric(unique_id), 
                                  root, data = redata23)
```

```{r}
summary(aster23_nb_plotfi_temp_diff,info.tol=1e-15)
```

#### Appendix 6: LRTs

Submodels:

Testing the significance of seeds:temp_diff_s

```{r}
aster23_nb_plotfi_temp_diff_1 <- aster(resp ~ varb+seeds:(I(temp_diff_s^2)+
                                                          poll_type+plot),
                                   pred, fam, famlist=famlist, varb,
                                   idvar=as.numeric(unique_id), 
                                   root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_temp_diff_1,aster23_nb_plotfi_temp_diff)
```

Testing the significance of seeds:I(temp_diff^2)

```{r}
aster23_nb_plotfi_temp_diff_2 <- aster(resp ~ varb+seeds:(temp_diff_s+
                                                          poll_type+plot), 
                                  pred, fam, famlist=famlist, varb,
                                  idvar=as.numeric(unique_id), 
                                  root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_temp_diff_2,aster23_nb_plotfi_temp_diff)
```

Testing the significance of seeds:poll_type

```{r}
aster23_nb_plotfi_temp_diff_3 <- aster(resp ~ varb+seeds:(temp_diff_s+
                                                          I(temp_diff_s^2)+
                                                          plot), 
                                  pred, fam, famlist=famlist, varb,
                                  idvar=as.numeric(unique_id), 
                                  root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_temp_diff_3,aster23_nb_plotfi_temp_diff)
```

Testing the significance of seeds:plot

```{r}
aster23_nb_plotfi_temp_diff_4 <- aster(resp ~ varb+seeds:(temp_diff_s+
                                                          I(temp_diff_s^2)+
                                                          poll_type), 
                                  pred, fam, famlist=famlist, varb,
                                  idvar=as.numeric(unique_id), 
                                  root, data = redata23)
```

```{r}
anova(aster23_nb_plotfi_temp_diff_4,aster23_nb_plotfi_temp_diff)
```

#### Predictions

```{r}
# Predictions calculated for values ranging from the 5th percentile to the 95th 
# percentile, to avoid making predictions based on very few observations
# NOT DONE HERE! Using min and max.

# make prediction df
aster23_nb_plotfi_temp_diff_predict1 <- data.frame(
  expand_grid(temp_diff_s = seq(from = as.numeric(range(redata23$temp_diff_s)[1]),
                           to = as.numeric(range(redata23$temp_diff_s)[2]),
                           length.out = 100)),
  #OK from here to below? Not sure...
  temp_s = 0, # scaled so mean=0
  surv = 1,
  fl_01 = 1,
  fr_01 = 1,
  nseed = 0,
  root = 1
)
aster23_nb_plotfi_temp_diff_predict1<-rbind(aster23_nb_plotfi_temp_diff_predict1%>%
                                         mutate(poll_type=as.factor("Crosspollinated")),
                                       aster23_nb_plotfi_temp_diff_predict1%>%
                                         mutate(poll_type=as.factor("Selfpollinated")))
aster23_nb_plotfi_temp_diff_predict1<-rbind(aster23_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[1])),
                               aster23_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[2])),
                               aster23_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[3])),
                               aster23_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[4])),
                               aster23_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[5])),
                               aster23_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[6])),
                               aster23_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[7])),
                               aster23_nb_plotfi_temp_diff_predict1%>%
                                 mutate(plot=as.factor(unique(redata23$plot)[8])))

# reshape data to long format
aster23_nb_plotfi_temp_diff_predict1_long <- reshape(as.data.frame(aster23_nb_plotfi_temp_diff_predict1), 
                                        varying = list(vars), direction = "long", 
                                        timevar = "varb", times = as.factor(vars), 
                                        v.names = "resp")

# add artifice
aster23_nb_plotfi_temp_diff_predict1_long$seeds <- as.integer(ifelse(
  aster23_nb_plotfi_temp_diff_predict1_long$varb == "nseed", 1, 0))

aster23_nb_plotfi_temp_diff_predict1.df <- aster23_nb_plotfi_temp_diff_predict1_long

# fixed model predictions
aster23_nb_plotfi_temp_diff.p1 <- predict(aster23_nb_plotfi_temp_diff,
                             newdata = aster23_nb_plotfi_temp_diff_predict1_long,
                             varvar = varb,idvar = id, root = root,
                             info.tol = 1e-15, 
                             se.fit = T)

aster23_nb_plotfi_temp_diff_predict1.df$fit <- aster23_nb_plotfi_temp_diff.p1$fit
aster23_nb_plotfi_temp_diff_predict1.df$se <- aster23_nb_plotfi_temp_diff.p1$se.fit

aster23_nb_plotfi_temp_diff_predict1.df.seeds <- aster23_nb_plotfi_temp_diff_predict1.df %>%
  filter(varb == "nseed")

aster23_nb_plotfi_temp_diff_predict1.df.seeds<-aster23_nb_plotfi_temp_diff_predict1.df.seeds%>%
  mutate(ymin=fit-se,ymax=fit+se)

aster23_nb_plotfi_temp_diff_predict1.df.seeds<-aster23_nb_plotfi_temp_diff_predict1.df.seeds%>%
  mutate(temp_diff=temp_diff_s*original_sd_temp_diff_23+original_mean_temp_diff_23)

aster23_nb_plotfi_temp_diff_predict1.df.seeds<-aster23_nb_plotfi_temp_diff_predict1.df.seeds%>%
  group_by(temp_diff)%>%
  summarise(fit=mean(fit),se=mean(se),ymin=mean(ymin),ymax=mean(ymax))
```

## Appendix 7

```{r}
App7<-grid.arrange(
  ggplot()+
    geom_point(data=data_transplants%>%filter(fruiting==1),
               aes(x=temp,y=n_tot_seed,color=year),size=1,alpha=0.2)+
    # 2022 temp
    geom_ribbon(data=aster22_nb_plotfi_temp_predict1.df.seeds,
                aes(x=temp,y=fit,ymin=ymin,ymax=ymax),
                fill="#F8766D",alpha=0.25)+
    geom_line(data=aster22_nb_plotfi_temp_predict1.df.seeds,
              aes(x=temp,y=fit),size=0.5,color="#F8766D")+
    # 2023 temp
    geom_ribbon(data=aster23_nb_plotfi_temp_predict1.df.seeds,
                aes(x=temp,y=fit,ymin=ymin,ymax=ymax),
                fill="#00BFC4",alpha=0.25)+
    geom_line(data=aster23_nb_plotfi_temp_predict1.df.seeds,
              aes(x=temp,y=fit),size=0.5,color="#00BFC4")+
    xlab("Temperature at planting site (ºC)")+ylab(NULL)+
    my_theme(),
  ggplot()+
    geom_point(data=data_transplants%>%filter(fruiting==1),
               aes(x=temp_diff,y=n_tot_seed,color=year),size=1,alpha=0.2)+
    # 2022 temp_diff
    geom_ribbon(data=aster22_nb_plotfi_temp_diff_predict1.df.seeds,
                aes(x=temp_diff,y=fit,ymin=ymin,ymax=ymax),
                fill="#F8766D",alpha=0.25)+
    geom_line(data=aster22_nb_plotfi_temp_diff_predict1.df.seeds,
              aes(x=temp_diff,y=fit),size=0.5,color="#F8766D")+
    # 2023 temp_diff
    geom_ribbon(data=aster23_nb_plotfi_temp_diff_predict1.df.seeds,
                aes(x=temp_diff,y=fit,ymin=ymin,ymax=ymax),
                fill="#00BFC4",alpha=0.25)+
    geom_line(data=aster23_nb_plotfi_temp_diff_predict1.df.seeds,
              aes(x=temp_diff,y=fit),size=0.5,color="#00BFC4")+
    xlab("Temperature difference (ºC)")+ylab(NULL)+
    my_theme(),
  ncol=2,left=textGrob("Predicted fitness",rot=90,
                       gp=gpar(fontsize=16,fontfamily="serif"),just="center"))
ggsave(file="output/figures/App7.tiff", plot=App7,width=7,height=3.5,dpi=300,
       compression="lzw")
```

```{r}
App7_nopoints<-grid.arrange(
  ggplot()+
    #geom_point(data=data_transplants%>%filter(fruiting==1),
    #           aes(x=temp,y=n_tot_seed,color=year),size=1.5,alpha=0.2)+
    # 2022 temp
    geom_ribbon(data=aster22_nb_plotfi_temp_predict1.df.seeds,
                aes(x=temp,y=fit,ymin=ymin,ymax=ymax),
                fill="#F8766D",alpha=0.25)+
    geom_line(data=aster22_nb_plotfi_temp_predict1.df.seeds,
              aes(x=temp,y=fit),size=0.5,color="#F8766D")+
    # 2023 temp
    geom_ribbon(data=aster23_nb_plotfi_temp_predict1.df.seeds,
                aes(x=temp,y=fit,ymin=ymin,ymax=ymax),
                fill="#00BFC4",alpha=0.25)+
    geom_line(data=aster23_nb_plotfi_temp_predict1.df.seeds,
              aes(x=temp,y=fit),size=0.5,color="#00BFC4")+
    xlab("Temperature at planting site (ºC)")+ylab(NULL)+
    my_theme()+scale_y_continuous(limits=c(0,85)),
  ggplot()+
    #geom_point(data=data_transplants%>%filter(fruiting==1),
    #           aes(x=temp_diff,y=n_tot_seed,color=year),size=1.5,alpha=0.2)+
    # 2022 temp_diff
    geom_ribbon(data=aster22_nb_plotfi_temp_diff_predict1.df.seeds,
                aes(x=temp_diff,y=fit,ymin=ymin,ymax=ymax),
                fill="#F8766D",alpha=0.25)+
    geom_line(data=aster22_nb_plotfi_temp_diff_predict1.df.seeds,
              aes(x=temp_diff,y=fit),size=0.5,color="#F8766D")+
    # 2023 temp_diff
    geom_ribbon(data=aster23_nb_plotfi_temp_diff_predict1.df.seeds,
                aes(x=temp_diff,y=fit,ymin=ymin,ymax=ymax),
                fill="#00BFC4",alpha=0.25)+
    geom_line(data=aster23_nb_plotfi_temp_diff_predict1.df.seeds,
              aes(x=temp_diff,y=fit),size=0.5,color="#00BFC4")+
    xlab("Temperature difference (ºC)")+ylab(NULL)+
    my_theme()+scale_y_continuous(limits=c(0,85)),
  ncol=2,left=textGrob("Predicted fitness",rot=90,
                       gp=gpar(fontsize=16,fontfamily="serif"),just="center"))
ggsave(file="output/figures/App7_nopoints.tiff",plot=App7_nopoints,
       width=7,height=3.5,dpi=300,compression="lzw")
```

# Session info

```{r}
sessionInfo()
```
