---
title: "Analyses Cerastium transplants"
subtitle: "Analyses meteorological data"
author : "Alicia Valdés"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: 4
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
options(tibble.width = Inf)
```

```{r load packages, include=FALSE}
library(tidyverse)
library(here)
library(readxl)
```

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
myPalette <- colorRampPalette(brewer.pal(11, "YlOrRd"))
```

# Read data

```{r}
meteo_data<-read_excel(
  here("data", "raw", "Hellisskarð weather station 2022-2023.xlsx"),
  sheet = "Weather observations")
```

# Data preparation

```{r}
meteo_data <- meteo_data %>%
  mutate(AR = as.factor(AR), MAN = as.factor(MAN)) %>%
  rename(year = AR, month = MAN, meanT = T, mean_max_T = TX,
         highest_T_month = TXX, mean_min_T = TN, lowest_T_month = TNN) %>%
  # Other variables are related to humidity, wind, air pressure and precipitation
  dplyr::select(year, month, meanT, mean_max_T, highest_T_month, mean_min_T,
         lowest_T_month)
```

Convert to long format:

```{r}
meteo_data_long <- meteo_data %>%
  pivot_longer(cols = meanT:lowest_T_month, names_to = "variable",
               values_to = "value")
```


# Plots

```{r}
ggplot(meteo_data_long,
       aes(x = month, y = value, color = variable, linetype = year,
           group = interaction(year, variable)))+
  geom_point() + geom_line() + my_theme_legend()
```

```{r}
summary(lm(highest_T_month ~ year,
           data = meteo_data %>% filter(month %in% c(4, 5, 6, 7))))
```

## OLD CODE from Maladaptive paper below


```{r}
ggplot(subset(meteo_data,day_of_year>=91&day_of_year<=212),
       aes(x=day_of_year,y=temp,color=year))+geom_line()+my_theme_legend()+
  ggtitle("April-July,temperature")+
  geom_vline(xintercept=91,linetype="dotted")+
  geom_vline(xintercept=121,linetype="dotted")+
  geom_vline(xintercept=152,linetype="dotted")+
  geom_vline(xintercept=182,linetype="dotted")+
  geom_vline(xintercept=212,linetype="dotted")
ggsave("output/figures/April-July_temperature.tiff",dpi=300)
```

```{r}
ggplot(subset(meteo_data,day_of_year>=91&day_of_year<=212),
       aes(x=day_of_year,y=max_temp_h,color=year))+geom_line()+my_theme_legend()+
  ggtitle("April-July,max temperature per hour")+
  geom_vline(xintercept=91,linetype="dotted")+
  geom_vline(xintercept=121,linetype="dotted")+
  geom_vline(xintercept=152,linetype="dotted")+
  geom_vline(xintercept=182,linetype="dotted")+
  geom_vline(xintercept=212,linetype="dotted")
ggsave("output/figures/April-July_max.tiff",dpi=300)
```

```{r}
ggplot(subset(meteo_data,day_of_year>=91&day_of_year<=212),
       aes(x=day_of_year,y=min_temp_h,color=year))+geom_line()+my_theme_legend()+
  ggtitle("April-July,min temperature per hour")+
  geom_vline(xintercept=91,linetype="dotted")+
  geom_vline(xintercept=121,linetype="dotted")+
  geom_vline(xintercept=152,linetype="dotted")+
  geom_vline(xintercept=182,linetype="dotted")+
  geom_vline(xintercept=212,linetype="dotted")
ggsave("output/figures/April-July_min.tiff",dpi=300)
```

# Compare montly temperatures

```{r}
meteo_data<-meteo_data%>%mutate(month=month(date))
```

```{r}
model_april_temp<-lm(temp~year,data=subset(meteo_data,month==4))
model_may_temp<-lm(temp~year,data=subset(meteo_data,month==5))
model_june_temp<-lm(temp~year,data=subset(meteo_data,month==6))
model_july_temp<-lm(temp~year,data=subset(meteo_data,month==7))
```

```{r}
summ(model_april_temp) # *, >2018
summ(model_may_temp) # *, <2018
summ(model_june_temp) # *, <2018
summ(model_july_temp) # p=0.06 (<2018)
```

```{r}
plot(ggpredict(model_april_temp))
plot(ggpredict(model_may_temp))
plot(ggpredict(model_june_temp))
plot(ggpredict(model_july_temp))
```

```{r}
model_april_max<-lm(max_temp_h~year,data=subset(meteo_data,month==4)) 
model_may_max<-lm(max_temp_h~year,data=subset(meteo_data,month==5)) 
model_june_max<-lm(max_temp_h~year,data=subset(meteo_data,month==6)) 
model_july_max<-lm(max_temp_h~year,data=subset(meteo_data,month==7)) 
```

```{r}
summ(model_april_max) # *, >2018
summ(model_may_max) # *, <2018
summ(model_june_max) # *, <2018
summ(model_july_max)# ns
```

```{r}
plot(ggpredict(model_april_max))
plot(ggpredict(model_may_max))
plot(ggpredict(model_june_max))
plot(ggpredict(model_july_max))
```

```{r}
model_april_min<-lm(min_temp_h~year,data=subset(meteo_data,month==4)) 
model_may_min<-lm(min_temp_h~year,data=subset(meteo_data,month==5)) 
model_june_min<-lm(min_temp_h~year,data=subset(meteo_data,month==6)) 
model_july_min<-lm(min_temp_h~year,data=subset(meteo_data,month==7)) 
```

```{r}
summ(model_april_min) # *, >2018
summ(model_may_min) # *, <2018
summ(model_june_min) # ns
summ(model_july_min)# p=0.07 (<2018)
```

```{r}
plot(ggpredict(model_april_min))
plot(ggpredict(model_may_min))
plot(ggpredict(model_june_min))
plot(ggpredict(model_july_min))
```















# Start, end and length of the growing season

## Start of the GS

Start date of the first 5-day period with mean daily temperatures over 5°C.

```{r}
# Check if there are missing dates
date_range <- seq(min(meteo_data$date), max(meteo_data$date), by = 1) 
date_range[!date_range %in% meteo_data$date] 
# [1] "2018-05-15" is missing

meteo_data<-meteo_data%>%add_row(station=31392,date=as.Date("2018-05-15"),
                                 year=as.factor(2018),month=5)%>%
  arrange(date)

# Make the rolling function
rolling_sum_5<-slidify(.f=sum,.period=5,.align="left",.partial=F)

# Apply rolling Function
meteo_data_GSstart<-meteo_data%>%mutate(temp_over_5=ifelse(temp>5,1,0))%>%
  group_by(year)%>%
  mutate(rolling_sum_5 = rolling_sum_5(temp_over_5))%>%
  filter(rolling_sum_5==5) %>% 
  filter(month==3|month==4|month==5|month==6)%>%
  slice(1)
meteo_data_GSstart
```

2017: May 2 (agrees with Bryndis calculation)
2018: June 2 (does not agree with Bryndis calculation)

## End of the GS

When the 10-day running mean falls below 5°C.

```{r}
# Make the rolling function
rolling_mean_10<-slidify(.f=mean,.period=10,.align="right",.partial=F)

# Apply rolling Function
meteo_data_GSend<-meteo_data%>%
  group_by(year)%>%
  mutate(rolling_mean_10 = rolling_mean_10(temp))%>%
  group_by(year)%>%
  filter(rolling_mean_10<5) %>% 
  filter(month>6)%>%
  slice(1)
meteo_data_GSend
```

2017: October 6
2018: September 20

## Length of the GS

```{r}
meteo_data_GSend$date-meteo_data_GSstart$date
```

2017: 157 days
2018: 110 days
(difference of 157-110=47 days longer in 2017)

# Number of days with freezing temperatures 

Number of days with minimum temperatures (min_temp_h) lower than 0.

## After GS start

```{r}
meteo_data%>%mutate(GSstart=as.Date(ifelse(year==2017,
                                           meteo_data_GSstart$date[1],
                                           meteo_data_GSstart$date[2])),
                    GSend=as.Date(ifelse(year==2017,
                                           meteo_data_GSend$date[1],
                                           meteo_data_GSend$date[2])))%>%
  filter(date>=GSstart&date<=GSend&min_temp_h<0)%>%
  group_by(year)%>%
  dplyr::select(date,min_temp_h)
```

3 days with freezing temperatures in spring 2017 after GS start.
No days with freezing temperatures in spring 2018 after GS start (2 days with freezing temperatures in autumn before GS end).

## In May

```{r}
meteo_data%>%filter(month==5&min_temp_h<0)%>%
  group_by(year)%>%
  dplyr::select(date,min_temp_h)%>%
  summarise(count=n())
```

## In June

```{r}
meteo_data%>%filter(month==6&min_temp_h<0)%>%
  group_by(year)%>%
  dplyr::select(date,min_temp_h)%>%
  summarise(count=n())
```

# Compare GS temperatures

```{r}
meteo_data%>%mutate(GSstart=as.Date(ifelse(year==2017,
                                           meteo_data_GSstart$date[1],
                                           meteo_data_GSstart$date[2])),
                    GSend=as.Date(ifelse(year==2017,
                                           meteo_data_GSend$date[1],
                                           meteo_data_GSend$date[2])))%>%
  filter(date>=GSstart&date<=GSend)%>%
  lm(temp~year,data=.)%>%summary()
meteo_data%>%mutate(GSstart=as.Date(ifelse(year==2017,
                                           meteo_data_GSstart$date[1],
                                           meteo_data_GSstart$date[2])),
                    GSend=as.Date(ifelse(year==2017,
                                           meteo_data_GSend$date[1],
                                           meteo_data_GSend$date[2])))%>%
  filter(date>=GSstart&date<=GSend)%>%
  lm(min_temp_h~year,data=.)%>%summary()
meteo_data%>%mutate(GSstart=as.Date(ifelse(year==2017,
                                           meteo_data_GSstart$date[1],
                                           meteo_data_GSstart$date[2])),
                    GSend=as.Date(ifelse(year==2017,
                                           meteo_data_GSend$date[1],
                                           meteo_data_GSend$date[2])))%>%
  filter(date>=GSstart&date<=GSend)%>%
  ggplot(aes(day_of_year,temp,color=year))+geom_line()+
  my_theme_legend()+ggtitle("Growing season, mean temperature")
meteo_data%>%mutate(GSstart=as.Date(ifelse(year==2017,
                                           meteo_data_GSstart$date[1],
                                           meteo_data_GSstart$date[2])),
                    GSend=as.Date(ifelse(year==2017,
                                           meteo_data_GSend$date[1],
                                           meteo_data_GSend$date[2])))%>%
  filter(date>=GSstart&date<=GSend)%>%
  ggplot(aes(day_of_year,min_temp_h,color=year))+geom_line()+
  my_theme_legend()+ggtitle("Growing season, min temperature")
```

No differences in GS temperatures.

# Session info

```{r include=FALSE}
sessionInfo()
```

